### 2024-02-02 index.qmd
update

---
title: Solar wind discontinuities spatial evolution in the outer heliosphere
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
  - name: Shi Chen
    orcid: 0000-0002-2582-7085
    affiliation: University of California, Los Angeles
affiliations:
  - id: ucla
    name: University of California, Los Angeles
execute: 
  eval: false
code-links:
  - repo
  - href: https://app.tana.inc?nodeid=Kmdzbyj-36NM
    text: Tana
  - href: https://www.overleaf.com/project/65b0a9c80a3616adf6e599e8
    text: Overleaf
abstract: We present a study of the spatial evolution of solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. We identify and analyze the properties of the discontinuities at different radial distances from the Sun. By differentiating the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance), we find that (1) the normalized occurrence rate of IDs drops with the radial distance from the Sun, following a $1/r$ law; (2) The thickness of IDs increases with the radial distance from the Sun, but after normalization to the ion inertial length, the thickness of IDs decreases; (3) The current intensity of IDs decreases with the radial distance from the Sun, but after normalization to the Alfven current, the current intensity of IDs increases.
---

# Introduction

Rapid variations in interplanetary magnetic fields, commonly recognized as solar wind magnetic discontinuities [@colburnDiscontinuitiesSolarWind1966], embody localized transient rotations or jumps of the magnetic field. Considered important for efficient plasma heating, they carry the most intense currents found in the solar wind and obserrved throughout the heliosphere from inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011]. Theoretical models [@lerchePropagationMagneticDisturbances1975, @medvedevDissipativeDynamicsCollisionless1997] and MHD simulations [@grecoIntermittentMHDStructures2008, @grecoStatisticalAnalysisDiscontinuities2009, @yangFormationRotationalDiscontinuities2015] suggest that the formation and destruction of discontinuities are closely related to the nonlinear dynamics of Alfvén waves and/or Alfvénic turbuluence. These nonlinear processes can create significant isolated disturbances to the otherwise adiabatic evolution of the solar wind flow (?Reference?) and host many processes, including magnetic reconnection (?Reference?) and Fermi acceleration of particles [@wentzelMotionMagneticDiscontinuities1964]. Moreover, they contribute significantly to the magnetic fluctuation spectra [@borovskyContributionStrongDiscontinuities2010] and can affect space weather [@tsurutaniReviewInterplanetaryDiscontinuities2011]. Therefore, the study of the spatial evolution of solar wind discontinuities is important for understanding the dynamics of the solar wind and test the local generation mechanism of discontinuities. However, previous study of solar wind discontinuities evolution [@sodingRadialLatitudinalDependencies2001] in the outer heliosphere were rarely in conjunction with measurements closer to the Sun. Moreover, the intervals in their study spread over multiple solar cycles though they are selected during solar activity minimum. Thus it is presently unclear whether their frequency and properties are the result of solar variability or due to the natural evolution of the discontinuities during their propagation and interaction with the ambient solar wind.

The goal of our paper is to investigate, statistically, discontinuities at different radial distance in the outer heliosphere to obtain a understanding of their formation and evolution. Our methodology involves integrating Juno measurements [@connerneyJunoMagneticField2017] between $1$ and $5$ AU with those at $1$ AU to distinguish temporal effect and spatial variations by examining the discontinuity characteristics at two radial distances simultaneously.

First, the missions, instruments and data used are listed. Then, we brieﬂy describe the method used to identify the discontinuities and the model used to estimate the plasma state at Juno location due to the lack of plasma data during its cruise phase.  Finally, we present the results of the spatial evolution of the discontinuities with respect to their characteristics (spatial scales and current density), and discuss the implications of our findings.

:::: {.content-visible when-format="html"}
::: {.callout-note collapse="false"}

## Related works

- @sodingRadialLatitudinalDependencies2001 studied the radial distribution of discontinuities in the solar wind between 0.3 and 19 AU during solar activity minimum.
- @liuCharacteristicsInterplanetaryDiscontinuities2021 studied the spatial evolution of the discontinuities from 0.13-0.9 AU, using measurements from the Parker Solar Probe.

[Typical rate of occurrence is about 50 per day at 1AU]{.aside}

::: {#fig-ids_distribution .column-page layout-ncol="2"}
[![Discontinuities between 0.3 and 19 AU](images/söding2001_fig4.png){fig-alt="Distribution of DD per day over r with a bin size of 0.1 AU and the standard deviations as error bars on logarithmic scale; TS criterion at the top and B at the bottom; darker lines represent intervals in the ecliptic which are used for the given fit and lighter ones are out of the ecliptic; right Voyager 2 in detail on linear scale."}](https://www.researchgate.net/figure/Distribution-of-DD-per-day-over-r-with-a-bin-size-of-01-AU-and-the-standard-deviations_fig4_29624789)

![Discontinuities between 0.13–0.9 AU](images/liu2021_fig2.png){fig-alt="Histograms as a function of heliocentric distance of (a) the number of events, (b) the PSP detection time, and (c)–(e) the occurrence rates of the total IDs, RDs and TDs. Panel (f) shows the ratio of RD occurrence rate to TD occurrence rate. The orange curves in panels (c) and (d) are the power-function fittings of ID and RD occurrence rates, with the coefficient of determination R2 being 0.983 and 0.985, respectively. The uncertainties of the power-law indices correspond to 95% confidence bounds."}

Spatial distribution of the discontinuities
:::
:::
::::

# Methods

- We use [@liuMagneticDiscontinuitiesSolar2022] method to identify IDs, which has better compatibility for the IDs with minor field changes.
- Then the minimum variance analysis is applied to each ID event to obtain the boundary normal (LMN) coordinate and extract IDs' features.

::: {#fig-examples .column-page layout-nrow="1"}

![](figures/examples/artemis_id_example)

![](figures/examples/stereo_id_example)

![](figures/examples/wind_id_example)

Examples of IDs from ARTEMIS, STEREO, and Wind

:::

:::: {.content-visible when-format="html"}
::: {.callout-note collapse="false"}
::: {#fig-ids_examples layout-ncol="2"}
![Current sheets in the heliosheath: Voyager 1, 2009](images/jgra21162-fig-0009.png){fig-alt="Hodograms produced by means of a minimum variance analysis for the current sheets associated with the sector boundaries in Figures 8 and 10. (b, d, and f) The value of the component of B along the minimum variance direction (Bm) as a function of time, N. (a, c, and e) The components of B along the intermediate variance direction (Bi) and maximum variance direction (BM)."}

![A rotational discontinuity (RD) detected at 0.126 AU](images/jgra57047-fig-0002-m.jpg){fig-alt="The properties of a rotational discontinuity (RD) detected at 0.126 AU on 28 January 2020. (a) The magnetic field magnitude and three components in Radial‐Tangential‐Normal (RTN) coordinate system, (b) the magnetic field in local magnetic normal (LMN) coordinate system, (c and d) the proton velocity in RTN and LMN coordinate systems, (e) the proton density, (f) the proton thermal pressure, the magnetic pressure and the sum, (g) the tangential components of the electric field in LMN coordinate system in the RD rest frame, (h) the position, orientation, and motion of the RD, (i–k) the distributions of BL, JM, and J · E across the RD, (l and m) the jump conditions across the RD. In (h), the cyan plane and arrow represent the RD and its motion vector, and the yellow curve shows the PSP orbit. In (j), both the colors and the curve correspond to the values of JM, and two vertical dashed lines indicate the area of $\vert {J}_{M}\vert > \frac{1}{3}\text{max}\left(\vert {J}_{M}\vert \right)$ whose width is defined as the half thickness of the discontinuity. In (m), the solid lines present the actual upstream and downstream velocities, while the dashed lines show the theoretical downstream velocity to satisfy the Walén relation."}

Discontinuities examples across the heliosphere
:::
:::
::::

## ID identification (limited feature extraction / anomaly detection)

Traditional methods for ID identiﬁcation, such as the criteria of

-   Burlaga & Ness (1969; B-criterion) : a directional change of the magnetic ﬁeld larger than 30° during 60 s
-   Tsurutani & Smith (1979; TS-criterion) : $|ΔB|/|B| \geq 0.5$ within 3 minutes

Mostly rely on magnetic ﬁeld variations with a certain time lag. B-criterion has, as its main condition.

In their methods, the IDs below the thresholds are artiﬁcially abandoned. Therefore, identiﬁcation criteria may affect the statistical results, and there is likely to be a discrepancy between the ﬁndings via B-criterion and TS- criterion.

Liu's method : The first two conditions guarantee that the field changes of the IDs identiﬁed are large enough to be distinguished from the stochastic fluctuations on magnetic fields, while the third is a supplementary condition to reduce the uncertainty of recognition.

$$ 
\textrm{Index}_1 = \frac{\sigma(\vec{B})}{Max(\sigma(\vec{B}_-),\sigma(\vec{B}_+))} 
$$

$$
\textrm{Index}_2 = \frac{\sigma(\vec{B}_- + \vec{B}_+)} {\sigma(\vec{B}_-) + \sigma(\vec{B}_+)}
$$

$$
\textrm{Index}_3 = \frac{| \Delta \vec{B} |}{|B_{bg}|}
$$

$$ 
\textrm{Index}_1 \ge 2, \textrm{Index}_2 \ge 1, \textrm{Index}_3 \ge 0.1 
$$

## Solar Wind Model

Sadly, JUNO does not provide plasma data during the cruise phase, so to estimate the plasma state we will use MHD model.

We are using [Michigan Solar WInd Model 2D (MSWIM2D)](http://csem.engin.umich.edu/MSWIM2D/), which models the solar wind propagation in 2D using the BATSRUS MHD solver. @keeblerMSWIM2DTwodimensionalOuter2022

Some key points about the model

-   Representing the solar wind in the ecliptic plane from 1 to 75 AU
-   2D MHD model, using the BATSRUS MHD solver
-   Inclusion of neutral hydrogen (important for the outer heliosphere)
-   Inner boundary is filled by time-shifting in situ data from multiple spacecraft

For model validation part, please see [JUNO Model Report](notebooks/20_model.ipynb).

# Conclusion

-   We have collected 5 years of solar wind discontinuities from JUNO, aMIS and STEREO.
-   We have developed a pipeline to identify solar wind discontinuities. (Modular, Performant, Scalable)
-   The normalized occurrence rate of IDs drops with the radial distance from the Sun, following $1/r$ law.
-   The thickness of IDs increases with the radial distance from the Sun, but after normalization to ion inertial length, the thickness of IDs decreases.
-   The current intensity of IDs decrease with the radial distance from the Sun, but after normalization to the Alfven current , the current intensity of IDs increases.

# Figures

![Juno orbit](figures/juno_orbit_white){#fig-orbit}

::: {#fig-examples .column-page layout-nrow="1"}

![](figures/examples/artemis_id_example)

![](figures/examples/stereo_id_example)

![](figures/examples/wind_id_example)

Examples of IDs from ARTEMIS, STEREO, and Wind

:::

![Near-Earth's solar wind plasma data during JUNO cruise phase](figures/omni_overview){#fig-overview}

Overview of the solar wind at 1 AU

[For code, see [noteboook](notebooks/20_omni_overview.ipynb).]{.aside}

![Model validation](figures/model/juno_model_validation_full){#fig-modelValidation}

![Effect of the time resolution on the discontinuity properties](figures/ts_effect){#fig-tsEffect}