### 2024-02-08 index.qmd
update

---
title: Solar wind discontinuities spatial evolution in the outer heliosphere
---

## Publications

- manuscript please see [here](_manuscript/index.html)
- AGU poster [here](presentations/AGU23_poster.qmd)

## Updates

::: {.column-screen layout-ncol="4"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

![](figures/new/thickness/thickness_k_r_dist)

![](figures/new/thickness/thickness_k_N1_r_dist)

New
:::

::: {.column-screen layout-ncol="4"}

![](figures/current/current_r_dist)

![](figures/current/current_N1_r_dist)

![](figures/thickness/thickness_k_r_dist)

![](figures/thickness/thickness_k_N1_r_dist)

Old
:::

::: {.column-screen layout-nrow="1"}

![Wind](figures/ts_method/wind)

![Juno](figures/ts_method/juno)

![Derivative method](figures/ts_method/derivative)

![Fit method](figures/ts_method/fit)

:::

![](figures/ts_method/all)

<!-- ::: {.column-screen layout-nrow="1"}

![Effect of the time resolution on the discontinuity properties](figures/ts_effect_fit_new)

![Effect of the time resolution on the discontinuity properties](figures/ts_effect)

::: -->

## Related works

-   @sodingRadialLatitudinalDependencies2001 studied the radial distribution of discontinuities in the solar wind between 0.3 and 19 AU during solar activity minimum.
-   @liuCharacteristicsInterplanetaryDiscontinuities2021 studied the spatial evolution of the discontinuities from 0.13-0.9 AU, using measurements from the Parker Solar Probe.

[Typical rate of occurrence is about 50 per day at 1AU]{.aside}

::: {#fig-ids_distribution .column-page layout-ncol="2"}
[![Discontinuities between 0.3 and 19 AU](images/söding2001_fig4.png){fig-alt="Distribution of DD per day over r with a bin size of 0.1 AU and the standard deviations as error bars on logarithmic scale; TS criterion at the top and B at the bottom; darker lines represent intervals in the ecliptic which are used for the given fit and lighter ones are out of the ecliptic; right Voyager 2 in detail on linear scale."}](https://www.researchgate.net/figure/Distribution-of-DD-per-day-over-r-with-a-bin-size-of-01-AU-and-the-standard-deviations_fig4_29624789)

![Discontinuities between 0.13–0.9 AU](images/liu2021_fig2.png){fig-alt="Histograms as a function of heliocentric distance of (a) the number of events, (b) the PSP detection time, and (c)–(e) the occurrence rates of the total IDs, RDs and TDs. Panel (f) shows the ratio of RD occurrence rate to TD occurrence rate. The orange curves in panels (c) and (d) are the power-function fittings of ID and RD occurrence rates, with the coefficient of determination R2 being 0.983 and 0.985, respectively. The uncertainties of the power-law indices correspond to 95% confidence bounds."}

Spatial distribution of the discontinuities
:::