### 2024-02-15 index.qmd
refactor mission pipeline

---
title: Solar wind discontinuities spatial evolution in the outer heliosphere
---

## Publications

- manuscript please see [here](_manuscript/index.html)
- AGU poster [here](presentations/AGU23_poster.qmd)

## Updates

::: {.column-screen layout-nrow="1"}

![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

New Fit Width
:::

::: {.column-screen layout-ncol="4"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

![](figures/new/thickness/thickness_k_r_dist)

![](figures/new/thickness/thickness_k_N1_r_dist)

New
:::

::: {.column-screen layout-ncol="4"}

![](figures/current/current_r_dist)

![](figures/current/current_N1_r_dist)

![](figures/thickness/thickness_k_r_dist)

![](figures/thickness/thickness_k_N1_r_dist)

Old
:::

::: {.column-screen layout-ncol="4"}

![Wind](figures/ts_method/wind)

![Juno](figures/ts_method/juno)

![Derivative method](figures/ts_method/derivative)

![Fit method](figures/ts_method/fit)

![](figures/ts_method/wind_fit)

![](figures/ts_method/all)

:::

<!-- ::: {.column-screen layout-nrow="1"}

![Effect of the time resolution on the discontinuity properties](figures/ts_effect_fit_new)

![Effect of the time resolution on the discontinuity properties](figures/ts_effect)

::: -->

## Related works

-   @sodingRadialLatitudinalDependencies2001 studied the radial distribution of discontinuities in the solar wind between 0.3 and 19 AU during solar activity minimum.
-   @liuCharacteristicsInterplanetaryDiscontinuities2021 studied the spatial evolution of the discontinuities from 0.13-0.9 AU, using measurements from the Parker Solar Probe.

[Typical rate of occurrence is about 50 per day at 1AU]{.aside}

::: {#fig-ids_distribution .column-page layout-ncol="2"}
[![Discontinuities between 0.3 and 19 AU](images/söding2001_fig4.png){fig-alt="Distribution of DD per day over r with a bin size of 0.1 AU and the standard deviations as error bars on logarithmic scale; TS criterion at the top and B at the bottom; darker lines represent intervals in the ecliptic which are used for the given fit and lighter ones are out of the ecliptic; right Voyager 2 in detail on linear scale."}](https://www.researchgate.net/figure/Distribution-of-DD-per-day-over-r-with-a-bin-size-of-01-AU-and-the-standard-deviations_fig4_29624789)

![Discontinuities between 0.13–0.9 AU](images/liu2021_fig2.png){fig-alt="Histograms as a function of heliocentric distance of (a) the number of events, (b) the PSP detection time, and (c)–(e) the occurrence rates of the total IDs, RDs and TDs. Panel (f) shows the ratio of RD occurrence rate to TD occurrence rate. The orange curves in panels (c) and (d) are the power-function fittings of ID and RD occurrence rates, with the coefficient of determination R2 being 0.983 and 0.985, respectively. The uncertainties of the power-law indices correspond to 95% confidence bounds."}

Spatial distribution of the discontinuities
:::

::: {.content-visible when-format="html"}
::: {.callout-note collapse="false"}
::: {#fig-ids_examples layout-ncol="2"}
![Current sheets in the heliosheath: Voyager 1, 2009](images/jgra21162-fig-0009.png){fig-alt="Hodograms produced by means of a minimum variance analysis for the current sheets associated with the sector boundaries in Figures 8 and 10. (b, d, and f) The value of the component of B along the minimum variance direction (Bm) as a function of time, N. (a, c, and e) The components of B along the intermediate variance direction (Bi) and maximum variance direction (BM)."}

![A rotational discontinuity (RD) detected at 0.126 AU](images/jgra57047-fig-0002-m.jpg){fig-alt="The properties of a rotational discontinuity (RD) detected at 0.126 AU on 28 January 2020. (a) The magnetic field magnitude and three components in Radial‐Tangential‐Normal (RTN) coordinate system, (b) the magnetic field in local magnetic normal (LMN) coordinate system, (c and d) the proton velocity in RTN and LMN coordinate systems, (e) the proton density, (f) the proton thermal pressure, the magnetic pressure and the sum, (g) the tangential components of the electric field in LMN coordinate system in the RD rest frame, (h) the position, orientation, and motion of the RD, (i–k) the distributions of BL, JM, and J · E across the RD, (l and m) the jump conditions across the RD. In (h), the cyan plane and arrow represent the RD and its motion vector, and the yellow curve shows the PSP orbit. In (j), both the colors and the curve correspond to the values of JM, and two vertical dashed lines indicate the area of $\\vert {J}_{M}\\vert > \\frac{1}{3}\\text{max}\\left(\\vert {J}_{M}\\vert \\right)$ whose width is defined as the half thickness of the discontinuity. In (m), the solid lines present the actual upstream and downstream velocities, while the dashed lines show the theoretical downstream velocity to satisfy the Walén relation."}

Discontinuities examples across the heliosphere
:::
:::
:::