### 2024-02-15 notebooks/_missions/index.qmd

---
title: Missions
---

This module includes helpful functions for working with mission (magnetic field and plasma) data. We currently support the following missions:

- [Parker Solar Probe](psp/index.qmd)
- [Juno](juno/index.ipynb)
- [Wind](wind/index.ipynb)
- [THEMIS/ARTEMIS](themis/index.ipynb)
- [STEREO](stereo/index.ipynb)

Note: `speasy` and `pyspedas` are not suitable for processing large datasets, although they are useful for downloading and processing small datasets.

- `speasy` cache file and data are not persistent
- `pyspedas` loads the entire dataset into memory