### 2024-05-13 notes.qmd

## Updates

### 03/15/2024

![Discontinuities in log-log scale](figures/l_j_fit_xscale_log)

![Same as above but with multiple time resolutions and time windows](figures/l_j_r_fit_xscale_log)

### 02/27/2024

::: {.column-screen}
::: {#fig-evolution-juno layout-nrow="1"}
![Evolution of plamsa property using fitting method](figures/l_j_r_fit.png)

![Evolution of plamsa property using derivative method](figures/l_j_r_der.png)

From top to bottom is Juno observation in 1,2,3,4,5 AU
:::
:::

### 02/26/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-comp >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-j-r-der >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-r-fit >}}
:::

### 02/25/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/archive/05_sea.ipynb#fig-fit >}}

{{< embed notebooks/archive/05_sea.ipynb#fig-fit-params >}}
:::

### 02/22/2024

::: {.column-screen layout-nrow="1"}
![First Year](figures/sea/sea_juno_first_year_high.png)

![Last Year](figures/sea/sea_juno_last_year_high.png)
:::

### 02/20/2024

### 02/19/2024

{{< embed notebooks/04_properties.ipynb#fig-jno-8hz-tau-20-der >}}

### 02/18/2024

![](figures/new/duration/duration_dist_ts_comparison)

#### Thickness
::: {.column-screen layout-nrow="1"}
![](figures/new/thickness/thickness_k_dist)

![](figures/new/thickness/thickness_k_dist-8hz)

![](figures/new/thickness/thickness_k_dist-8hz-tau_20)
:::

#### Current Density

::: {.column-screen layout-nrow="1"}
![](figures/new/current/current_k_dist)

![](figures/new/current/current_k_dist-8hz)

![](figures/new/current/current_k_dist-8hz-tau_20)

:::

### 02/17/2024

::: {.column-screen layout-nrow="1"}

![](figures/new/duration/duration_dist)

![](figures/new/thickness/thickness_k_fit_dist)

![](figures/new/current/current_k_dist)
:::

::: {.column-screen layout-nrow="1"}

![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

New Fit Width
:::

::: {.column-screen layout-ncol="4"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

![](figures/new/thickness/thickness_k_r_dist)

![](figures/new/thickness/thickness_k_N1_r_dist)

New
:::

::: {.column-screen layout-ncol="4"}

![](figures/current/current_r_dist)

![](figures/current/current_N1_r_dist)

![](figures/thickness/thickness_k_r_dist)

![](figures/thickness/thickness_k_N1_r_dist)

Old
:::

::: {.column-screen layout-ncol="4"}

![Wind](figures/ts_method/wind)

![Juno](figures/ts_method/juno)

![Derivative method](figures/ts_method/derivative)

![Fit method](figures/ts_method/fit)

![](figures/ts_method/wind_fit)

![](figures/ts_method/all)

:::

<!-- ::: {.column-screen layout-nrow="1"}

![Effect of the time resolution on the discontinuity properties](figures/ts_effect_fit_new)

![Effect of the time resolution on the discontinuity properties](figures/ts_effect)

::: -->