### 2024-05-13 presentations/rMeeting.qmd

---
title: "Current Sheets in the Solar Wind: JUNO Observations in the Outer Heliosphere"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
format:
  html: default
  pptx: default
slide-number: true
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms
-->

# References