### 2024-05-24 presentations/.rMeeting.qmd

---
title: "Current Sheets in the Solar Wind: JUNO Observations in the Outer Heliosphere"
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
format:
  html: default
  pptx: default
slide-number: true
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms
-->

## Motivation

Studying the radial evolution of solar wind discontinuities from observations of PSP (0.1 AU – 1 AU) and JUNO (1AU – 5AU) with Earth-orbiting missions (ARTEMIS, Wind) during aligned intervals.

## Methods

![Time resolution effect](../figures/method/time_res_effect.png)

![Tau effect](../figures/method/tau_effect.png)

## References