### 2024-10-09 notes.qmd
update

---
title: Notes
---

# Data

`updated_events_JNO_method=fit_tau=0:01:00_ts=0:00:01.arrow` and `events.JNO.fit.ts_1.00s_tau_60s.arrow` are the same data for backward compatibility.

`updated_events_Wind_tr=20110825-20160630_method=fit_tau=0:01:00_ts=0:00:00.090909.arrow` and `events.Wind.fit.ts_0.09s_tau_60s.arrow`

# Updates

### 09/23/2024

::: {.column-page}
::: {#fig-ocr-v layout-nrow="1"}

![Wind](figures/ocr/ocr_wind_daily_v.svg)

![](images/söding2001_fig4.png)

Distribution of daily averages of DD per day over v with the fit of an linear function.
:::
:::

### 03/15/2024

![Discontinuities in log-log scale](figures/l_j_fit_xscale_log)

![Same as above but with multiple time resolutions and time windows](figures/l_j_r_fit_xscale_log)

### 02/27/2024

::: {.column-screen}
::: {#fig-evolution-juno layout-nrow="1"}
![Evolution of plamsa property using fitting method](figures/l_j_r_fit.png)

![Evolution of plamsa property using derivative method](figures/l_j_r_der.png)

From top to bottom is Juno observation in 1,2,3,4,5 AU
:::
:::

<!-- 
### 02/26/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-comp >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-j-r-der >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-r-fit >}}
:::

### 02/25/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/archive/05_sea.ipynb#fig-fit >}}

{{< embed notebooks/archive/05_sea.ipynb#fig-fit-params >}}
::: 
-->

### 02/22/2024

::: {.column-screen layout-nrow="1"}
![First Year](figures/sea/sea_juno_first_year_high.png)

![Last Year](figures/sea/sea_juno_last_year_high.png)
:::

### 02/20/2024

### 02/19/2024

<!-- {{< embed notebooks/04_properties.ipynb#fig-jno-8hz-tau-20-der >}} -->

### 02/18/2024

![](figures/new/duration/duration_dist_ts_comparison)

#### Thickness
::: {.column-screen layout-nrow="1"}
![](figures/new/thickness/thickness_k_dist)

![](figures/new/thickness/thickness_k_dist-8hz)

![](figures/new/thickness/thickness_k_dist-8hz-tau_20)
:::

#### Current Density

::: {.column-screen layout-nrow="1"}
![](figures/new/current/current_k_dist)

![](figures/new/current/current_k_dist-8hz)

![](figures/new/current/current_k_dist-8hz-tau_20)

:::

### 02/17/2024

::: {.column-screen layout-nrow="1"}

![](figures/new/duration/duration_dist)

![](figures/new/thickness/thickness_k_fit_dist)

![](figures/new/current/current_k_dist)
:::

::: {.column-screen layout-nrow="1"}

![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

New Fit Width
:::

::: {.column-screen layout-ncol="4"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

![](figures/new/thickness/thickness_k_r_dist)

![](figures/new/thickness/thickness_k_N1_r_dist)

New
:::

::: {.column-screen layout-ncol="4"}

![](figures/current/current_r_dist)

![](figures/current/current_N1_r_dist)

![](figures/thickness/thickness_k_r_dist)

![](figures/thickness/thickness_k_N1_r_dist)

Old
:::

::: {.column-screen layout-ncol="4"}

![Wind](figures/ts_method/wind)

![Juno](figures/ts_method/juno)

![Derivative method](figures/ts_method/derivative)

![Fit method](figures/ts_method/fit)

![](figures/ts_method/wind_fit)

![](figures/ts_method/all)

:::

## Solar Wind Model

Sadly, JUNO does not provide plasma data during the cruise phase, so to estimate the plasma state we will use MHD model.

We are using [Michigan Solar WInd Model 2D (MSWIM2D)](http://csem.engin.umich.edu/MSWIM2D/), which models the solar wind propagation in 2D using the BATSRUS MHD solver. @keeblerMSWIM2DTwodimensionalOuter2022

Some key points about the model

-   Representing the solar wind in the ecliptic plane from 1 to 75 AU
-   2D MHD model, using the BATSRUS MHD solver
-   Inclusion of neutral hydrogen (important for the outer heliosphere)
-   Inner boundary is filled by time-shifting in situ data from multiple spacecraft

For model validation part, please see [JUNO Model Report](notebooks/20_model.ipynb).

<!-- 
::: {.column-screen layout-nrow="1"}
![Effect of the time resolution on the discontinuity properties](figures/ts_effect_fit_new)

![Effect of the time resolution on the discontinuity properties](figures/ts_effect)
:::
-->

# TODOs

Science part

-   Analysis
    -   [ ] Re-calculate the occurrence rate of IDs
    -   [x] Distribution of $δB/B$ over radius (related to the alfvenicity, identification standard about neighboring fluctuations)
    -   [x] $\theta_{n,v}$ distribution
    -   [ ] Correlation between plasma beta and thickness / current density

    -   [ ] Check Datagap
    -   [ ] Check `ARTEMIS-B` data in different states (solar wind, magnetosheath, magnetotail, moon wake)
    -   [ ] Distribution of \|B\| over radius
    -   [ ] JUNO from 2012-09\~2012-10 lack of IDS and extreme large thickness
    -   [ ] Wind data
    -   [ ] Add error bar
    -   [ ] Validate the effects of calibrate candidate duration

    -   ~~Contribution of discontinuities to the power spectrum~~
    -   ~~Validate model density with `Voyager`~~
-   Identifaction
    -   [ ] Ensemble forest?
    -   [ ] Smoothing is important?
    -   [ ] Check change point algorithm
-   Visualize data gaps
-   Features
    -   [ ] Thickness in N direction
    -   [ ] Use high resolution data for feature extraction
-   Compare with other methods of identifying IDs
    -   [ ] Verify with other methods of identifying IDs
-   [x] Incorporate solar wind propagation model
    -   [x] Verify with solar wind propagation model
        -   [x] Coordinate transformation

# Archived Figures

::: {#fig-current layout-nrow="1"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

**a,** Current density of IDs as a function of the radial distance from the Sun. **b,** Current density of IDs normalized to the Alfven current as a function of the radial distance from the Sun.
:::

::: {#fig-thickness layout-nrow="1"}
![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

**a,** Thickness of IDs as a function of the radial distance from the Sun. **b,** Thickness of IDs normalized to the ion inertial length as a function of the radial distance from the Sun. 
:::

![](figures/l_j_fit_taus)

![](figures/l_j_fit_taus_xscale_log)

::: {.column-screen-inset layout-nrow="1"}
![](figures/thickness/thickness_k_dist)

![](figures/thickness/thickness_k_N1_dist)

![](figures/thickness/thickness_k_N1_r_dist)
:::

::: {.column-screen-inset layout-nrow="1"}

![](figures/current/current_k_dist)

![](figures/current/current_k_N1_dist)

:::