### 2024-11-04 notebooks/01_results_julia.qmd
update

---
title: Results (Julia)
---

```{julia}
using DrWatson
using Discontinuity
@quickactivate
include("../src/main.jl");
M = var_mapping()
```

```{julia}
wind_df = load_wind();
jno_df = load_jno();
```

## Check the discontinuities properties with time and radial distance

```{julia}
maps1 = [M.l_log M.l_norm_log; M.j_log M.j_norm_log]
maps2 = [M.density_log M.B_log; beta_map M.bm0_log];
```

```{julia}
temp_plot(layer::Layer) = m -> plot_dist(layer, m; datalimits=datalimits_f)
temp_plot(df::DataFrame; color=year_map) = temp_plot(data(df) * mapping(color=color))

let df = wind_df, fnames = ["wind_distribution_time", "wind_sw_paramters"]
    fs = temp_plot(df).([maps1, maps2])
    easy_save.(fnames, fs)
    display.(fs)
end
```

```{julia}
# let df = j_events_taus
let df = jno_df, color = r_map, fnames = ["juno_distribution_r", "juno_sw_paramters"]
    fs = temp_plot(df; color).([maps1, maps2])
    easy_save.(fnames, fs)
    display.(fs)
end
```