### 2024-11-10 article.qmd
update: add discussion about Bn, PVI comparison, references to mission and clarification...

---
title: Spatial evolution of solar wind discontinuities in the outer heliosphere
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
affiliations:
  - id: ucla
    name: University of California, Los Angeles
execute: 
  eval: false
code-links:
  - href: ../
    text: Website
  - repo
  - href: https://www.overleaf.com/project/65b0a9c80a3616adf6e599e8
    text: Overleaf
abstract: We present a study of the spatial evolution of solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. We identify and analyze the properties of the discontinuities at different radial distances from the Sun. By differentiating the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance), we find that (1) the normalized occurrence rate of interplanetary discontinuities (IDs) decreases with radial distance from the Sun, following a $1/r$ relationship; (2) The thickness of IDs increases with radial distance, but when normalized to the ion inertial length, the thickness decreases; (3) The current density of IDs decreases with radial distance, but when normalized to the Alfvén current, the current density of IDs increases; (4) The distribution of the normalized thickness and current density of IDs remain constant over time at 1 AU on a yearly scale.
---

# Plain Language Summary

- **What I did:** I studied the solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. I identified and analyzed the properties of the discontinuities at different radial distances from the Sun. I differentiated the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance).

- **What I found:** The normalized occurrence rate of IDs drops with the radial distance from the Sun, following a $1/r$ law. The thickness of IDs increases with the radial distance from the Sun, but after normalization to the ion inertial length, the thickness of IDs decreases. The current intensity of IDs decreases with the radial distance from the Sun, but after normalization to the Alfven current, the current intensity of IDs increases.

- **What it means:** The results of this study provide a better understanding of the solar wind discontinuities in the outer heliosphere and their spatial evolution. This information is important for understanding the dynamics of the solar wind and testing the local generation mechanism of discontinuities.

# Introduction

TODO: annihilation mechanisms and related references
TODO: References to stabliity and time scales

Rapid variations in interplanetary magnetic fields, commonly recognized as solar wind magnetic discontinuities [@colburnDiscontinuitiesSolarWind1966], embody localized transient rotations or jumps of the magnetic field. They are microscale structures observed throughout the heliosphere, from inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011]. Despite being discovered over half a century ago, the specific origins and circumstances of their formation remain subjects of ongoing debate. 
Historically, it has been believed that discontinuities originate inside 1 AU, probably near or at the Sun, and are then convected passively outward by the solar wind [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001; @borovskyFluxTubeTexture2008]. However, the good agreement observed in the waiting-time analysis between solar wind data and magnetohydrodynamic (MHD) turbulence simulations [@grecoStatisticalAnalysisDiscontinuities2009] indicates that a significant portion of these discontinuities could be associated with flux-tube boundaries. These structures are likely intermittent, arising spontaneously and forming rapidly and locally within MHD turbulence. 
The annihilation mechanisms for these discontinuities are also uncertain. These discontinuities may be unstable features and some could rapidly wane or decay [@neugebauerTangentialDiscontinuitiesSolar1986]. Conversely, they can be very stable and persist over time intervals comparable to the convection time or the solar rotation period.
Further complicating this picture is the presence of different types of discontinuities in interplanetary space—specifically tangential and rotational discontinuities, which are classically regarded as static solutions to the MHD equations. These distinct types may arise from different sources and may evolve in unique ways. Investigating discontinuities across large radial distances has the potential to address these fundamental questions regarding their origin and evolution.

Discontinuities are not only an interesting topic in themselves but are also considered crucial for understanding the dynamics of the solar wind and cosmic rays. Theoretical models [@lerchePropagationMagneticDisturbances1975; @medvedevDissipativeDynamicsCollisionless1997] and MHD simulations [@grecoIntermittentMHDStructures2008; @yangFormationRotationalDiscontinuities2015] suggest that the formation and destruction of discontinuities are closely related to the nonlinear dynamics of Alfvén waves and Alfvénic turbulence. These nonlinear processes can create significant disturbances to the otherwise adiabatic evolution of the solar wind flow. 
These theoretical predictions are supported by spacecraft observations, which have revealed that the solar wind magnetic field follows the Parker model of the heliospheric current sheet only on average. Localized transient currents, which can be significantly more intense than those predicted by the Parker model, are carried by various discontinuities observed as strong variations in magnetic field components [@colburnDiscontinuitiesSolarWind1966; @burlagaMicroscaleStructuresInterplanetary1968; @turnerOrientationsRotationalTangential1971]. These discontinuities contribute significantly to the magnetic fluctuation spectra [@borovskyContributionStrongDiscontinuities2010] and play an important role in energy dissipation processes such as plasma heating [@osmanIntermittencyLocalHeating2012] and particle acceleration [@wentzelMotionMagneticDiscontinuities1964; @dmitrukTestParticleEnergization2004; @tesseinAssociationSuprathermalParticles2013]. 
More specifically, because tangential discontinuities act as boundaries separating different types of plasma and often involve oppositely directed magnetic fields on either side, they are potential regions where magnetic reconnection or merging can occur. Additionally, rotational discontinuities, being nonlinear large-amplitude Alfvén waves, are expected to significantly contribute to the diffusion coefficient for cosmic rays. Finally, these discontinuities can produce geomagnetic effects [@tsurutaniReviewInterplanetaryDiscontinuities2011] and impact space weather due to abrupt changes in external (solar wind) conditions.

In this study, we present a statistical analysis of discontinuities observed from Earth’s orbit to Jupiter, utilizing data collected by the Juno spacecraft [@boltonJunoMission2017] during its cruise phases (2011-2016), with its orbits shown in panel (a) of @fig-overview. The primary goal of this paper is to investigate the radial dependence of the occurrence rate and other properties of discontinuities, in order to gain insights into their evolution and origin. Continuous measurements at 1 AU from missions such as Wind, ARTEMIS, and STEREO have been utilized to distinguish temporal effects from spatial variations by examining the characteristics of discontinuities at two radial distances simultaneously. This distinction is crucial because temporal variations could easily be misinterpreted as spatial variations [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. Due to the potential misidentification of the normal direction of the magnetic field when using the minimum variance analysis (MVA) method [@hausmanDeterminingNatureOrientation2004; @liuFailuresMinimumVariance2023], we do not differentiate between the types of discontinuities, whether they are tangential, rotational, or either [@neugebauerReexaminationRotationalTangential1984].

Previous studies on the radial dependencies of solar wind discontinuities have been conducted by several authors [e.g., @marianiVariationsOccurrenceRate1973; @tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001]. However, the data used in these studies were rarely obtained in conjunction with measurements at fixed radial distances, and the methods employed to identify discontinuities were not robust for situations involving weak magnetic fields and long durations encountered at large radial distances. Additionally, some measurements at different radial distances were significantly separated in time, with limited time ranges covered [@sodingRadialLatitudinalDependencies2001]. Consequently, there remains ambiguity in interpreting the results of these studies.

The paper is structured as follows. First, we outline the missions, instruments, and data utilized in this study. Next, we provide a brief description of the method used to identify discontinuities and the model employed to estimate the plasma state at the Juno location, necessitated by the absence of plasma data during its cruise phase. In the subsequent sections, we present the results concerning the spatial variation of discontinuities, focusing on their occurrence rate and characteristics, such as spatial scales and current density. Finally, we discuss the implications of our findings in the context of the origin and evolution of these discontinuities in the solar wind.

# Dataset, models, and methods

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase of the JUNO mission, the sunspot number reached its peak, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

::: {#fig-overview}

![](figures/fig_overview)

Overview. **a,** Juno's orbit during its cruise phase (2011-2016). **b,** Difference in heliographic longitude between Juno and 1-AU missions. **c,** monthly and smoothed sunspot numbers. **d-g,** solar wind plasma density and speed from Near-Earth (OMNI) and STEREO-A missions.

:::

Further analysis of @fig-overview highlights the significance of the heliographic longitudinal difference between the Juno mission and Earth as well as the STEREO-A mission. Specifically, when there is a substantial longitudinal difference between Juno and Earth, the difference between Juno and STEREO-A tends to be minimal, and vice versa. This longitudinal discrepancy between Earth and STEREO-A ensures comprehensive coverage of the plasma en route to Juno.

## Dataset

In this study, we analyze Juno observations during its cruise phase from August 25, 2011, to July 30, 2016, spanning approximately 1,800 days. Data collected from missions such as STEREO, ARTEMIS, and WIND complement Juno’s observations, providing a broader perspective on heliospheric conditions and enabling a multi-point analysis of discontinuity properties [@velliUnderstandingOriginsHeliosphere2020]. Specifically, we utilize magnetic fields from Juno [@boltonJunoMission2017], measured by the Juno Fluxgate Magnetometer (MAG) [@connerneyJunoMagneticField2017], alongside ion bulk velocity and plasma density with hourly resolution from a solar wind model at Juno’s location (described below). For ARTEMIS [@angelopoulosARTEMISMission2011], we use magnetic field data from the Fluxgate Magnetometer [@austerTHEMISFluxgateMagnetometer2008] and ion bulk velocity and plasma density derived from the Electrostatic Analyzer [@mcfaddenTHEMISESAPlasma2009]. For STEREO[@kaiserSTEREOMissionIntroduction2008], we incorporate magnetic fields from the magnetic field experiment on IMPACT [@acuñaSTEREOIMPACTMagnetic2008; @luhmannSTEREOIMPACTInvestigation2008] and ion bulk velocity and plasma density from PLASTIC [@galvinPlasmaSuprathermalIon2008]. For WIND, we include magnetic fields from the Magnetic Field Investigation (MFI) [@leppingWINDMagneticField1995] and ion bulk velocity and plasma density from the Solar Wind Experiment (SWE) [@ogilvieSWEComprehensivePlasma1995]. The table below summarizes the time resolution of magnetic field and plasma data for each mission. These synergistic observations are expected to enhance our understanding of solar wind discontinuities, including their radial distribution and evolution.

| Mission             | δt(B)    | δt(plasma) |
|---------------------|----------|------------|
| Juno                | 1 Hz     | 1 hour     |
| ARTEMIS             | 5 Hz     | 0.25 Hz    |
| WIND                | 11 Hz    | 1 Hz       |
| STEREO              | 8 Hz     | 1 min      |

However, since Juno does not provide direct plasma measurements, we will estimate the spatial scale (thickness) of discontinuities using solar wind speed derived from solar wind propagation models. Specifically, we will employ the Two-Dimensional Outer Heliosphere Solar Wind Modeling (MSWIM2D) [@keeblerMSWIM2DTwodimensionalOuter2022] to determine the ion bulk velocity ($v$) and plasma density ($n$) at Juno’s location. This model, which utilizes the BATSRUS MHD solver[@tothAdaptiveNumericalAlgorithms2012], simulates the propagation of solar wind from 1 to 75 astronomical units (AU) in the ecliptic plane, effectively covering the region pertinent to our study. The MSWIM2D model provides output data with an hourly time resolution as shown in @fig-model. The comparison of magnetic field magnitudes from MSWIM2D with those measured by Juno, after averaging to the same time resolution, reveals a strong correlation, confirming the model’s reliability.

::: {#fig-model}

![](figures/model/juno_model_validation_full)

**a,** Magnetic field magnitude from MSWIM2D and Juno. **b-c,** Plasma speed and density from MSWIM2D model. **d,** Juno radial distance from the Sun.
:::

## Methods

We employ Liu’s method [@liuMagneticDiscontinuitiesSolar2022] to identify discontinuities in the solar wind. This approach provides better compatibility for discontinuities with small magnetic field changes and demonstrates greater robustness for conditions in the outer heliosphere compared to traditional methods, namely the B-criterion [@burlagaDirectionalDiscontinuitiesInterplanetary1969] and the TS-criterion [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. For each sampling time $t$, we define three intervals: the pre-interval $[-1,-1/2]\cdot T+t$, the middle interval $[-1/2,1/2]\cdot T+t$, and the post-interval $[1/2,1]\cdot T+t$, where $T$ represents time lags. The magnetic field timeseries in these three intervals are labeled as ${\mathbf B}_-$, ${\mathbf B}_0$, and ${\mathbf B}_+$, respectively. To identify a discontinuity, the following three conditions must be met:

$$
\begin{aligned}
\sigma({\mathbf B}_0) &> 2\max\left(\sigma({\mathbf B}_-), \sigma({\mathbf B}_+)\right) \\
\sigma\left({\mathbf B}_-+{\mathbf B}_+\right) &>\sigma({\mathbf B}_-)+\sigma({\mathbf B}_+) \\
|\Delta {\mathbf B}| &>|{\mathbf B}_{bg}|/10
\end{aligned}
$$

Here, $\sigma$ denotes the standard deviation, ${\mathbf B}_{bg}$ represents the background magnetic field magnitude, and $\Delta {\mathbf B}={\mathbf B}(t+T/2)-{\mathbf B}(t-T/2)$. The first two conditions ensure that the field changes are distinguishable from stochastic fluctuations, while the third condition serves as a supplementary measure to reduce recognition uncertainty.

Notably, this method shares certain similarities with the Partial Variance of Increments (PVI) approach for identifying coherent structures, as both calculate indices based on standard deviations (with this method focusing on the standard deviation of the magnetic field, and PVI on magnetic field increments). However, this method is better suited for discontinuity identification by comparing data within neighboring time windows rather than across a correlation scale (an unknown quantity that may vary with radial distance and solar wind conditions). As a result, this approach effectively excludes wave-like coherent structures, eliminating the need for manual inspection and bypassing the necessity of grouping continuous clusters of high-PVI points as in PVI analysis.

Because the discontinuities may have different spatial scale at different radial distances from the Sun, we repeat the search procedure with time lags $T = 10, 20, 30, 40, 50, 60$ s, and combine the identified ID events together to establish the event list.
Discontinuities with spatial scales larger than $60$ s are not considered in this study, however it could be argued that they are not significant in the outer heliosphere if they follow the trend with decreasing number of IDs with increasing spatial scale. @fig-junoDistribution also indicates that the spatial scale of IDs does not vary significantly with the radial distance from the Sun.

<!-- TODO: -->

More than ? IDs are identified in the solar wind data from Juno, with ? IDs identified in 5 AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum or maximum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal (LMN) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent functions to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data to obtain the thickness and the current density of the discontinuity.

$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta T}\right) + c_l
$${#eq-fit}

where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity, $\Delta T$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

Previous studies have indicated that the minimum variance direction obtained through Minimum Variance Analysis (MVA) is frequently unreliable. Improved accuracy can be achieved by either using the cross product of the magnetic field measured at the boundaries as the normal direction [@knetterFourpointDiscontinuityObservations2004] or by enforcing conditions such as $Δ|B|/|B| > 0.05$ or $ω > 60°$ [@liuFailuresMinimumVariance2023]. The former approach requires a small normal component in the magnetic field, while the latter approach limits the dataset to a subset that satisfies these conditions. In our analysis, we applied both methods and observed no significant differences in the results.

<!-- TODO: use both methods to determine the normal direction of the discontinuity and compare the results. -->

@fig-examples shows several examples of solar wind discontinuities detected by different spacecraft.

<!-- - We use [@liuMagneticDiscontinuitiesSolar2022] method to identify IDs, which has better compatibility for the IDs with minor field changes.
- Then the minimum variance analysis is applied to each ID event to obtain the boundary normal (LMN) coordinate and extract IDs' features. -->

::: {#fig-examples}
![Examples of IDs from Juno, ARTEMIS, STEREO, and Wind](figures/fig_examples)

Examples of solar wind discontinuities observed by various spacecraft at different heliocentric distances and times. Panels (a)–(c) display data from Juno at 1 AU (a), 3 AU (b), and 5 AU (c), showing magnetic field components $B_l$ (blue), $B_n$ (red), $B_m$ (green), and the magnetic field magnitude $B_{\text{total}}$ (black). Panels (d)–(f) show magnetic field and ion speed observations from other spacecraft: ARTEMIS on March 27, 2012 (d), STEREO on June 28, 2016 (e), and Wind on August 26, 2011 (f).
:::

# Results

<!-- The basic approach of this proposal is to use solar wind measurements at 1AU (STEREO, ARTEMIS, WIND) to compare with Juno and PSP measurements and distinguish effects of solar wind temporal variations and effects of spatial (with the radial distance from the Sun) variations of discontinuity occurrence rate and characteristics. The example of such comparison for the occurrence rate is shown in @fig-rate, where we plot number of discontinuities measured per day by different spacecraft missions (for the same temporal resolution of magnetic field data and the same criteria of discontinuity determination). The radial distance of Juno for 2011-2016 is shown in @fig-model, and the number of discontinuities measured by Juno per day coincides with the discontinuity number measured by STEREO, WIND, and ARTEMIS, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance (with time after $\sim 2013$), as Juno moves from $1$ AU to $5$ AU. We will use the similar comparison for discontinuity characteristics and occurrence rate derived for PSP and Juno.  -->

## Occurrence rate

The occurrence rate of IDs is investigated with regards of the radial distance from the Sun for Juno, ARTEMIS, STEREO, and Wind. Number of discontinuities measured per day by different spacecraft missions (for the same temporal resolution of magnetic field data and the same criteria of discontinuity determination) is shown in @fig-rate. Because data is not continuously available and ARTEMIS spacecrafts will only be exposed in the solar wind for a relative short time during its orbits around the Moon (the Moon would be in the magnetosphere for a long time as it orbits the earth), the occurrence rate (numbder of IDs per day) is corrected with respect to data gap and the time the mission spent in the pristine solar wind. 
The occurrence rate of IDs observed by Juno is consistent with the occurrence rate observed by 1AU missions, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance as Juno moves from $1$ AU to $5$ AU (corresponding to time after $\sim 2013$, the radial distance of Juno for 2011-2016 is shown in @fig-model)

The rate of occurrence depends linearly on the solar wind velocity [@sodingRadialLatitudinalDependencies2001], however the plasma velocity data in Juno mission is not available during its cruise phase. Therefore, we do not normalize the occurrence rate with respect to the solar wind speed. This influence of the solar wind speed on the occurrence rate might be mitigated by the fact all the spacecrafts are in the ecliptic plane, and the solar wind speed is not expected to change significantly with the radial distance from the Sun beyond $1$ AU. We normalize the occurrence rate from Juno to the occurrence rate from 1AU missions to account for the effect of solar wind structures on the occurrence rate. 
The normalized occurrence rate is shown in panel (b) of @fig-rate. The occurrence rate of IDs drops with the radial distance from the Sun, following $1/r$ law. So the decrease of the occurrence rate with increasing radial distance from the Sun is expected to be a real effect of the radial distance from the Sun and this confirms the former results [@sodingRadialLatitudinalDependencies2001].

::: {#fig-rate layout-nrow="1"}

![](figures/ocr/ocr_time_cleaned)

![](figures/ocr/ocr_r_N2_cleaned_simplified.pdf)

The number of discontinuities measured by Juno per day coincides with the discontinuity number measured by STEREO, WIND, and ARTEMIS, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance (with time after $\sim 2013$), as Juno moves from $1$ AU to $5$ AU. We will use the similar comparison for discontinuity characteristics and occurrence rate derived for PSP and Juno. The radial distance of Juno for 2011-2016 is shown in @fig-model.
:::

## Current density and thickness

The current density $J_m$ across a discontinuity is determined by

$$
J_m = - \frac{1}{μ_0 V_n} \frac{d B_l}{d t},
$$ {#eq-j}

where $μ_0$ is the vacuum permeability and $V_n$ is the normal component of the proton flow velocity. For fitting method, the derivative of the magnetic field could be estimated as $\max(d B_l / d t) = B_{l0} / \Delta T$. This estimate is more robust than the direct calculation of the derivative of the magnetic field, because the fitting method is less sensitive to the noise in the data and the time resolution of the data (see discussion in the appendix).

<!--  The proton flow velocity $v$ is expressed as $v = v_{sw} - v_{s/c} + U$, where $v_{sw}$ is the solar wind velocity, $v_{s/c}$ is the spacecraft velocity, and $U$  denotes the propagation speed of the discontinuities. In our case, the latter two velocities are negligible. -->

The thickness $d$ of a discontinuity is determined by

$$
d = V_n \Delta T
$$

where $\Delta T$ is the transition time derived from the fitting method.

To understand the relationship between the discontinuity and the local plasma, we normalize the thickness of the discontinuity by the initial ion length and the current density by the Alfven current density $J_A = e N_p V_A$ (corresponding to the drift between protons and electrons of local Alfén speed $V_A$). Figure @fig-junoDistribution presents the statistical distribution of the (un)normalized thickness and current density of IDs grouped by the radial distance from the Sun collected by Juno. 
Panel (a,b) shows that the thickness increases with radial distance. However, after normalization, the thickness slightly decreases, with the most probable value around 4-2 $d_i$. With increasing radial distance, larger fraction of IDs have thickness less than 1 $d_i$, becombing sub-proton scale structures.  Panel (c,d) shows that the current density decreases with increasing radial distance. However, after normalization, it slightly increases, with the most probable value around 0.05-0.15 $J_A$. It should be noted that for the normalized thickness and current density, the change is not significant compared to the spread of the distribution.

The distribution of the thickness and current density of IDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) during the five year of Juno cruise phase are shown in @fig-windDistribution. The distribution are grouped by the year of observation. Panel (a,b) shows that both the thickness and normalized thickness distribution of IDs remains almost constant with the year of observation. Most of the discontinuities have thickness around 350-3500 km, with the most probable value around 1000 km. Panel (c,d) shows that the current density slightly increases with the year of observation, with the most probable value around 5 $nA m^{-2}$, in consistence with previous studies [@vaskoKineticscaleCurrentSheets2022]. However, after normalization, the current density distribution remains identical across the five years. The change in the current density could be explained by the change in the solar wind properties with the solar cycle, as shown in @fig-swParameters. Density and magnetic field asscoiated with the discontinuities increases with the solar cycle approaching the solar minimum, which could lead to the increase of the current density of the discontinuities with fixed normalized current density.

::: {#fig-junoDistribution}
![](figures/juno_distribution_r)

Distribution of various properties of IDs observed by Juno grouped by the radial distance from the Sun. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

::: {#fig-windDistribution}
![](figures/wind_distribution_time)

Distribution of various properties of IDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) grouped by the year of observation. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

::: {#fig-swParameters}
![](figures/wind_sw_paramters)

Solar wind parameters associated with the IDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) grouped by the year of observation. Panel (a) solar wind density, (b) magnetic field, (c) fitted magnetic field $A$ in @eq-fit.
:::

# Discussion

## Spatial variations of discontinuity properties

TODO:

## Dependences of occurrence rate

The observed increase in the average waiting time between discontinuities with radial distance prompts us to explore different hypotheses to explain this phenomenon. At any given radial distance, discontinuities may originate from two primary sources: those generated near the Sun and convected outward by the solar wind, and those formed locally. Since the solar wind speed is significantly higher than the Alfvén speed—which characterizes the propagation speed of rotational discontinuities (RDs)—the fraction of inward-propagating discontinuities is expected to be negligible. Additionally, when observations are made by a single spacecraft, some discontinuities may be missed due to unfavorable positioning or limitations in detection methods.

First, we consider the limitations of our current detection methods. An increase in the duration of discontinuities could render some of them unidentifiable if their duration becomes comparable to the time interval used in our analysis. Although we observe a slight increase in the average duration of discontinuities with radial distance, this increase is relatively small. The overall population of discontinuities is still predominantly characterized by those with short durations (see Figure X in the appendix). Since our method normalizes the change in the magnetic field by the surrounding magnetic field, a decrease in magnetic field variation with radial distance alone does not necessitate non-identification. Notably, our results differ from previous studies [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979], which reported typical durations of 2 to 3 seconds at 1 AU and 10 to 30 seconds at 5 AU. In our analysis, however, the observed durations remain below 10 seconds at 5 AU, suggesting that our method may be more effective at identifying shorter-duration events than those reported in prior studies.

If the majority of distant discontinuities are generated locally, and assuming that the primary conditions for their generation do not change significantly with radial distance, we would expect them to be uniformly distributed throughout the hemisphere. Consequently, the occurrence rate should be independent of radial distance. To test this hypothesis, we examined the relationship between waiting time and plasma beta. However, even at a 99% statistical significance level, we could not reject the null hypothesis that these two variables are independent. Therefore, if discontinuities are primarily generated locally, there must be other factors influencing their generation rates.

Alternatively, we consider the hypothesis that the majority of discontinuities are generated close to the Sun. If we assume that discontinuities are one-dimensional, infinitely extended structures, then without annihilation, all discontinuities passing through one location would also pass through other radial distances. However, a more realistic assumption is that discontinuities are three-dimensional curved surfaces with lengths and widths proportional to their thickness. In this scenario, the occurrence rate observed by a single spacecraft does not represent the total number of discontinuities within a given shell. Since our spacecraft travels within the ecliptic plane and the normal direction of most discontinuities lies within this plane, it is necessary to project the discontinuities onto a circular cross-section. The projected lengths should then be normalized by the circumference of the circle to estimate the total number of discontinuities within the cylindrical shell. The normalization factor is  $d \sin\theta_{n,v} / r$ , where $d$ is the thickness of the discontinuity, $\theta_{n,v}$ is the angle between the normal direction of the discontinuity and the solar wind velocity, and  $r$ is the radial distance from the Sun (the result is shown in the appendix). Accounting for this factor reduces the differences in the normalized occurrence rates observed by different spacecraft. However, we cannot rule out the possibility that some discontinuities may be annihilated as they convect outward with the solar wind. This annihilation process would result in a lower number of discontinuities observed by Juno at larger radial distances. Therefore, under this hypothesis, the annihilation process must play a significant role. (However, if we disregard changes in the thickness and only account for the factor of the circumference of the circle, the normalized discontinuities observed by Juno would be consistent with those observed at 1 AU.)

![](figures/ocr/ocr_normalized_by_r)

# Conclusion

We have collected 5 years of solar wind discontinuities from JUNO, ARTEMIS and STEREO. The results can be summarized as follows:

1. The normalized occurrence rate of IDs drops with the radial distance from the Sun, following $1/r$ law.
2. The thickness of IDs increases with the radial distance from the Sun, but after normalization to ion inertial length, the thickness of IDs slightly decreases. The thickness distribution of IDs remains almost constant with the year of observation.
3. The current density of IDs decrease with the radial distance from the Sun, but after normalization to the Alfven current , the current density of IDs slightly increases. The current density distribution of IDs slightly increases with the year of observation, but after normalization, the current density distribution remains identical across the five years, thus no effect of the long-term (i.e. solar cycle) temporal variations.

# Acknowledgements

Sunspot data from the World Data Center SILSO, Royal Observatory of Belgium, Brussels. ARTEMIS solar wind time interval data from the OMNI database. ARTEMIS solar wind time interval data from OMNIWEB, link: [omniweb.gsfc.nasa.gov/ftpbrowser/themis_b_sw.txt](https://omniweb.gsfc.nasa.gov/ftpbrowser/themis_b_sw.txt).

# Appendix

## Effect of the time resolution on the discontinuity properties

The transition time $\Delta T$ is determined in one second resolution data for all our missions to reduce the system error. (Note that highest time resolution of the data is not needed for the transition time determination by fitting method).

![Effect of the time resolution on the discontinuity properties](figures/ts_effect){#fig-tsEffect}