### 2024-11-21 notes.qmd
doc: update

---
title: Notes
---

# Data

`updated_events_JNO_method=fit_tau=0:01:00_ts=0:00:01.arrow` and `events.JNO.fit.ts_1.00s_tau_60s.arrow` are the same data for backward compatibility.

`updated_events_Wind_tr=20110825-20160630_method=fit_tau=0:01:00_ts=0:00:00.090909.arrow` and `events.Wind.fit.ts_0.09s_tau_60s.arrow`

## Notation

Data Level

-   l0: unprocessed

-   l1: cleaned data, fill null value, add useful columns

-   l2: time-averaged data

### Columns naming conventions

-   `v_{x,y,z}` or `sw_vel_{X,Y,Z}`: solar wind plasma speed in the *ANY* coordinate system, in units of $km/s$

    -   `sw_vel_{r,t,n}`: solar wind plasma speed in the RTN coordinate system, in units of $km/s$
    -   `sw_vel_gse_{x,y,z}`: solar wind plasma speed in the GSE coordinate system, in units of $km/s$
    -   `sw_vel_lmn_{x,y,z}`: solar wind plasma speed in the LMN coordinate system, in units of $km/s$
        -   `v_l` or `sw_vel_l`: abbreviation for `sw_vel_lmn_1`
        -   `v_mn` or `sw_vel_mn` (deprecated)

-   `radial_distance`: radial distance of the spacecraft, in units of $AU$

-   `sw_elevation`: solar wind elevation angle, in units of $\degree$

-   `sw_azimuth`: solar wind azimuth angle, in units of $\degree$

-   `b_rtn_{x,y,z}` or `b_{r,t,n}`: magnetic field in the RTN coordinate system

-   `b_gse_{x,y,z}`: magnetic field in the GSE coordinate system

-   `model_b_{r,t,n}`: modelled magnetic field in the RTN coordinate system

-   `state` : *1* for *solar wind*, *0* for *non-solar wind*

## Solar Wind Model

Sadly, JUNO does not provide plasma data during the cruise phase, so we use MHD model to estimate the plasma state.

[Michigan Solar WInd Model 2D (MSWIM2D)](http://csem.engin.umich.edu/MSWIM2D/) models the solar wind propagation in 2D using the BATSRUS MHD solver [@keeblerMSWIM2DTwodimensionalOuter2022].

Some key points about the model

-   Representing the solar wind in the ecliptic plane from 1 to 75 AU
-   2D MHD model, using the BATSRUS MHD solver
-   Inclusion of neutral hydrogen (important for the outer heliosphere)
-   Inner boundary is filled by time-shifting in situ data from multiple spacecraft

For model validation part, please see [JUNO Model Report](notebooks/20_model.ipynb).

# Updates

### 09/23/2024

::: {.column-page}
::: {#fig-ocr-v layout-nrow="1"}

![Wind](figures/ocr/ocr_wind_daily_v.svg)

![](images/söding2001_fig4.png)

Distribution of daily averages of DD per day over v with the fit of an linear function.
:::
:::

### 03/15/2024

![Discontinuities in log-log scale](figures/l_j_fit_xscale_log)

![Same as above but with multiple time resolutions and time windows](figures/l_j_r_fit_xscale_log)

### 02/27/2024

::: {.column-screen}
::: {#fig-evolution-juno layout-nrow="1"}
![Evolution of plamsa property using fitting method](figures/l_j_r_fit.png)

![Evolution of plamsa property using derivative method](figures/l_j_r_der.png)

From top to bottom is Juno observation in 1,2,3,4,5 AU
:::
:::

<!-- 
### 02/26/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-comp >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-j-r-der >}}
:::

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/01_results_julia.ipynb#fig-l-r-fit >}}
:::

### 02/25/2024

::: {.column-screen layout-nrow="1"}
{{< embed notebooks/archive/05_sea.ipynb#fig-fit >}}

{{< embed notebooks/archive/05_sea.ipynb#fig-fit-params >}}
::: 
-->

### 02/22/2024

::: {.column-screen layout-nrow="1"}
![First Year](figures/sea/sea_juno_first_year_high.png)

![Last Year](figures/sea/sea_juno_last_year_high.png)
:::

### 02/20/2024

### 02/19/2024

<!-- {{< embed notebooks/04_properties.ipynb#fig-jno-8hz-tau-20-der >}} -->

### 02/18/2024

![](figures/new/duration/duration_dist_ts_comparison)

#### Thickness
::: {.column-screen layout-nrow="1"}
![](figures/new/thickness/thickness_k_dist)

![](figures/new/thickness/thickness_k_dist-8hz)

![](figures/new/thickness/thickness_k_dist-8hz-tau_20)
:::

#### Current Density

::: {.column-screen layout-nrow="1"}
![](figures/new/current/current_k_dist)

![](figures/new/current/current_k_dist-8hz)

![](figures/new/current/current_k_dist-8hz-tau_20)

:::

### 02/17/2024

::: {.column-screen layout-nrow="1"}

![](figures/new/duration/duration_dist)

![](figures/new/thickness/thickness_k_fit_dist)

![](figures/new/current/current_k_dist)
:::

::: {.column-screen layout-nrow="1"}

![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

New Fit Width
:::

::: {.column-screen layout-ncol="4"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

![](figures/new/thickness/thickness_k_r_dist)

![](figures/new/thickness/thickness_k_N1_r_dist)

New
:::

::: {.column-screen layout-ncol="4"}

![](figures/current/current_r_dist)

![](figures/current/current_N1_r_dist)

![](figures/thickness/thickness_k_r_dist)

![](figures/thickness/thickness_k_N1_r_dist)

Old
:::

::: {.column-screen layout-ncol="4"}

![Wind](figures/ts_method/wind)

![Juno](figures/ts_method/juno)

![Derivative method](figures/ts_method/derivative)

![Fit method](figures/ts_method/fit)

![](figures/ts_method/wind_fit)

![](figures/ts_method/all)

:::

<!-- 
::: {.column-screen layout-nrow="1"}
![Effect of the time resolution on the discontinuity properties](figures/ts_effect_fit_new)

![Effect of the time resolution on the discontinuity properties](figures/ts_effect)
:::
-->

# TODOs

Science part

-   Analysis
    -   [ ] Texts about waiting time distribution
    -   [ ] Re-calculate the occurrence rate of IDs
    -   [x] Distribution of $δB/B$ over radius (related to the alfvenicity, identification standard about neighboring fluctuations)
    -   [x] $\theta_{n,v}$ distribution
    -   [ ] Correlation between plasma beta and thickness / current density

    -   [ ] Check datagap and visualize the data gap
    -   [x] Check `ARTEMIS-B` data in different states (solar wind, magnetosheath, magnetotail, moon wake)
    -   [ ] JUNO from 2012-09\~2012-10 lack of IDS and extreme large thickness
    -   [x] Wind data

    -   ~~Validate the effects of calibrate candidate duration~~
    -   ~~Contribution of discontinuities to the power spectrum~~
    -   ~~Validate model density with `Voyager`~~
-   Identifaction
    -   [ ] Ensemble forest?
    -   [x] Smoothing is important?
    -   [ ] Check change point algorithm
-   Compare with other methods of identifying IDs
    -   [ ] Verify with other methods of identifying IDs
-   [x] Incorporate solar wind propagation model
    -   [x] Verify with solar wind propagation model
        -   [x] Coordinate transformation

# Archived Figures

::: {#fig-current layout-nrow="1"}
![](figures/new/current/current_k_r_dist)

![](figures/new/current/current_k_N1_r_dist)

**a,** Current density of IDs as a function of the radial distance from the Sun. **b,** Current density of IDs normalized to the Alfven current as a function of the radial distance from the Sun.
:::

::: {#fig-thickness layout-nrow="1"}
![](figures/new/thickness/thickness_k_fit_r_dist)

![](figures/new/thickness/thickness_k_fit_N1_r_dist)

**a,** Thickness of IDs as a function of the radial distance from the Sun. **b,** Thickness of IDs normalized to the ion inertial length as a function of the radial distance from the Sun. 
:::

![](figures/l_j_fit_taus)

![](figures/l_j_fit_taus_xscale_log)

::: {.column-screen-inset layout-nrow="1"}
![](figures/thickness/thickness_k_dist)

![](figures/thickness/thickness_k_N1_dist)

![](figures/thickness/thickness_k_N1_r_dist)
:::

::: {.column-screen-inset layout-nrow="1"}

![](figures/current/current_k_dist)

![](figures/current/current_k_N1_dist)

:::