### 2024-12-29 article.qmd
docs: incorporate Anton's revision

---
title: "Solar wind discontinuities in the outer heliosphere: spatial distribution between 1 and 5 AU"
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
  - name: Ivan Vasko
    affiliation: University of Texas at Dallas
affiliations:
  - id: ucla
    name: University of California, Los Angeles
execute: 
  eval: false
code-links:
  - href: ../
    text: Website
  - repo
  - href: https://www.overleaf.com/project/65b0a9c80a3616adf6e599e8
    text: Overleaf
abstract: Solar wind discontinuities (SWDs; also known as current sheets) are kinetic plasma structures contributing significantly to the spectrum of magnetic field fluctuations and to solar wind heating. Properties of SWDs varies significantly with the radial distance away from the Sun, and investigation of this radial evolution is important for revealing of SWD origin and their contribution to solar wind plasma processes. In this study we utilize quite unique dataset of Juno spacecraft observations of SWDs during cruise phase, when Juno was traveling from Earth ($\sim 1$AU) to Jupiter ($\sim 5$AU). We combine Juno measurements with Wind, ARTEMIS, and STEREO measurements of SWDs around $\sim 1$AU, and this combination allows us to separate temporal (with solar activity) and spatial (with radial distance) variations of SWD properties. We find that (1) the normalized to $1$AU occurrence rate of SWDs decreases with radial distance from the Sun, following a $1/r$ relationship; (2) the SWD thickness increases with radial distance, but when normalized to the ion inertial length, the thickness remains almost unchanged within $[1,5]$AU; (3) the SWD current density decreases with radial distance, but when normalized to the Alfvén current, the current density remains almost unchanged within $[1,5]$AU; (4) the distribution of SWD normalized thickness and current density remain constant over time (five year period of Juno cruise phase) at $1$ AU. We discuss these results in context of theories of SWD origin and dynamics.
# We present a study of the spatial evolution of solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. We identify and analyze the properties of the discontinuities at different radial distances from the Sun. By differentiating the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance), we find that (1) the normalized occurrence rate of interplanetary discontinuities (SWDs) decreases with radial distance from the Sun, following a $1/r$ relationship; (2) The thickness of SWDs increases with radial distance, but when normalized to the ion inertial length, the thickness decreases; (3) The current density of SWDs decreases with radial distance, but when normalized to the Alfvén current, the current density of SWDs increases; (4) The distribution of the normalized thickness and current density of SWDs remain constant over time at 1 AU on a yearly scale.
---

# Plain Language Summary

<!-- - **What I did:** I studied the solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. I identified and analyzed the properties of the discontinuities at different radial distances from the Sun. I differentiated the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance).
- **What I found:** The normalized occurrence rate of SWDs drops with the radial distance from the Sun, following a $1/r$ law. The thickness of SWDs increases with the radial distance from the Sun, but after normalization to the ion inertial length, the thickness of SWDs decreases. The current intensity of SWDs decreases with the radial distance from the Sun, but after normalization to the Alfven current, the current intensity of SWDs increases.
- **What it means:** The results of this study provide a better understanding of the solar wind discontinuities in the outer heliosphere and their spatial evolution. This information is important for understanding the dynamics of the solar wind and testing the local generation mechanism of discontinuities. -->
# Introduction

<!-- TODO: annihilation mechanisms and related references
TODO: References to stabliity and time scales -->

Rapid variations in the interplanetary magnetic field, commonly recognized as solar wind magnetic discontinuities (SWDs) [@colburnDiscontinuitiesSolarWind1966], embody localized transient rotations or jumps of the magnetic field. They are micro (ion kinetic) scale plasma structures observed throughout the heliosphere, from the inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011]. Despite being discovered over half a century ago, the specific origins and circumstances of their formation remain subjects of ongoing debate \cite<see discussion in>[and references therein]{Vasko22:apjl,Vasko24:jgr}. 
Historically, SWDs have been believed to originate inside 1 AU, probably near or at the Sun, and are then passively convected outward by the solar wind [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001; @borovskyFluxTubeTexture2008]. However, the good agreement observed in the waiting-time analysis between solar wind data and magnetohydrodynamic (MHD) turbulence simulations [@grecoStatisticalAnalysisDiscontinuities2009] indicates that a significant portion of these discontinuities could be associated with flux-tube boundaries, intermittent structures, arising spontaneously and forming rapidly and locally within MHD turbulence. 
The annihilation mechanisms for these discontinuities are also uncertain. These discontinuities may be unstable features and some could rapidly wane or decay [@neugebauerTangentialDiscontinuitiesSolar1986]. Magnetic reconnection of SWDs is also considered a natural process that destroys them \cite{Gosling12,Wilson16:cs,else}. Conversely, SWDs can be very stable and persist over time intervals comparable to the convection time or the solar rotation period.
Further complicating this picture is the presence of different types of discontinuities in interplanetary space -— specifically tangential and rotational discontinuities \cite<see discussion in>{Neugebauer06,Artemyev19:jgr:solarwind,wangSolarWindCurrent2024}, which are classically regarded as static solutions to the MHD equations [@hudsonDiscontinuitiesAnisotropicPlasma1970]. These distinct types may arise from different sources and may evolve in unique ways. Investigating discontinuities across large radial distances has the potential to address these fundamental questions regarding their origin and evolution.

Discontinuities are not only interesting because they are nonlinear plasma structures widely observed in the heliosphere,  but are also considered crucial for understanding the dynamics of the solar wind plasma and cosmic rays. Theoretical models [@lerchePropagationMagneticDisturbances1975; @medvedevDissipativeDynamicsCollisionless1997] and MHD simulations [@grecoIntermittentMHDStructures2008; @yangFormationRotationalDiscontinuities2015] suggest that the formation and destruction of discontinuities are closely related to the nonlinear dynamics of Alfvén waves and Alfvénic turbulence. These nonlinear processes can create significant disturbances to the otherwise adiabatic evolution of the solar wind flow [@matteiniIonKineticsSolar2012; @tsurutaniReviewAlfvenicTurbulence2018].
These theoretical predictions are supported by spacecraft observations, which have revealed that the solar wind magnetic field follows the Parker model of the heliospheric current sheet only on average [@svirzhevskyHeliosphericMagneticField2021]. Localized transient currents carried by SWDs and significantly more intense than those predicted by the Parker model, are observed as strong variations in magnetic field components [@colburnDiscontinuitiesSolarWind1966; @burlagaMicroscaleStructuresInterplanetary1968; @turnerOrientationsRotationalTangential1971]. SWDs contribute significantly to magnetic fluctuation spectra [@borovskyContributionStrongDiscontinuities2010; @vaskoKineticScaleCurrentSheets2024] and play an important role in energy dissipation processes such as plasma heating [@osmanIntermittencyLocalHeating2012] and particle acceleration [@wentzelMotionMagneticDiscontinuities1964; @dmitrukTestParticleEnergization2004; @tesseinAssociationSuprathermalParticles2013]. 
More specifically, because tangential discontinuities act as boundaries separating different types of plasma and often involve oppositely directed magnetic fields on either side, they are potential regions where magnetic reconnection or merging can occur \cite<see reviews by>{Gosling12,Paschmann13}. Additionally, rotational discontinuities, being nonlinear large-amplitude Alfvén waves, are expected to contribute significantly to the scattering of energetic particles [@artemyevSuperfastIonScattering2020].
<!-- Finally, these discontinuities can produce geomagnetic effects [@tsurutaniReviewInterplanetaryDiscontinuities2011] and impact space weather due to abrupt changes in external (solar wind) conditions. -->

Investigation of the evolution of SWD properties with the radial distance (or alternatively, investigation of properties of SWD generation at different radial distances) has mostly been limited to comparison of SWD statistics collected by different missions during different intervals of solar activity [e.g., @marianiVariationsOccurrenceRate1973; @tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001]. However, the data used in these studies were rarely obtained in conjunction with measurements at fixed radial distances, and the methods employed to identify discontinuities were not robust for situations involving weak magnetic fields and long durations encountered at large radial distances. Additionally, some measurements at different radial distances were significantly separated in time, with limited time ranges covered [@sodingRadialLatitudinalDependencies2001]. Consequently, there remains an ambiguity in interpreting the results of these studies.
In this study, we present a statistical analysis of discontinuities observed from Earth's orbit ($\sim 1$ AU) to Jupiter orbit ($\sim 5$ AU) by a single mission. We utilize data collected by the Juno spacecraft [@boltonJunoMission2017] during its cruise phases (2011-2016), with its orbits shown in panel (a) of @fig-overview. The primary goal of this study is to investigate the radial dependence of the SWD occurrence rate and properties, in order to gain insights into their evolution and origin. Continuous measurements of SWDs at 1 AU from missions such as Wind [@acunaGlobalGeospaceScience1995], ARTEMIS [@angelopoulosARTEMISMission2011], and STEREO [@kaiserSTEREOMissionIntroduction2008] have been utilized to distinguish temporal effects from spatial variations by examining the characteristics of discontinuities at two radial distances ($1$ AU and at Juno location) simultaneously. This distinction is crucial because temporal variations could easily be misinterpreted as spatial variations [see discussion in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. Due to the potential misidentification of the SWD normal plane with single spacecraft [see discussion in @hausmanDeterminingNatureOrientation2004; @liuFailuresMinimumVariance2023; @wangSolarWindCurrent2024], we do not differentiate between the types of discontinuities, whether they are tangential, rotational, or either [@neugebauerReexaminationRotationalTangential1984].

The paper is structured as follows. First, we outline the missions, instruments, and data utilized in this study (@sec-dataset). Next, we provide a brief description of the method used to identify discontinuities and the model employed to estimate the plasma state at the Juno location, necessitated by the absence of plasma data during its cruise phase (@sec-methods). In subsequent @sec-results, we present the results concerning the spatial variation of discontinuities, focusing on their occurrence rate and characteristics, such as spatial scales and current density. Finally, we discuss the implications of our findings in the context of the origin and evolution of these discontinuities in the solar wind (@sec-dependence).

# Dataset, models, and methods

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase (2011) of the Juno mission [@boltonJunoMission2017], the sunspot number peaked, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase in 2016, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

::: {#fig-overview}

![](figures/fig_overview)

Overview. **a,** Juno's orbit during its cruise phase (2011-2016). **b,** Difference in heliographic longitude between Juno and 1-AU missions. **c,** monthly and smoothed sunspot numbers. **d-g,** solar wind plasma density and speed from Near-Earth (OMNI) and STEREO-A missions.

:::

Further analysis of @fig-overview highlights the significance of the heliographic longitudinal difference between the Juno mission and Earth as well as the STEREO-A mission [@kaiserSTEREOMissionIntroduction2008]. Specifically, when there is a substantial longitudinal difference between Juno and Earth, the difference between Juno and STEREO-A tends to be minimal, and vice versa. This longitudinal discrepancy between Earth and STEREO-A ensures comprehensive coverage of the plasma en route to Juno.

## Dataset {#sec-dataset}

In this study, we analyze Juno observations during its cruise phase from August 25, 2011, to July 30, 2016, spanning approximately 1,800 days. Data collected from missions such as STEREO, ARTEMIS, and WIND complement Juno’s observations, providing a broader perspective on heliospheric conditions and enabling a multi-point analysis of discontinuity properties [@velliUnderstandingOriginsHeliosphere2020]. Specifically, we utilize magnetic fields from Juno [@boltonJunoMission2017], measured by the Juno Fluxgate Magnetometer (MAG) [@connerneyJunoMagneticField2017]. For ARTEMIS [@angelopoulosARTEMISMission2011], we use magnetic field data from the Fluxgate Magnetometer [@austerTHEMISFluxgateMagnetometer2008] and ion bulk velocity and plasma density derived from the Electrostatic Analyzer [@mcfaddenTHEMISESAPlasma2009]. For STEREO[@kaiserSTEREOMissionIntroduction2008], we incorporate magnetic fields from the magnetic field experiment on IMPACT [@acunaSTEREOIMPACTMagnetic2008; @luhmannSTEREOIMPACTInvestigation2008] and ion bulk velocity and plasma density from PLASTIC [@galvinPlasmaSuprathermalIon2008]. For WIND [@acunaGlobalGeospaceScience1995], we include magnetic fields from the Magnetic Field Investigation (MFI) [@leppingWINDMagneticField1995] and ion bulk velocity and plasma density from the Solar Wind Experiment (SWE) [@ogilvieSWEComprehensivePlasma1995]. The table below summarizes the time resolution of magnetic field and plasma data for each mission. These synergistic observations are expected to enhance our understanding of solar wind discontinuities, including their radial distribution and evolution.

| Mission             | δt(B)    | δt(plasma) |
|---------------------|----------|------------|
| Juno                | 1 Hz     | 1 hour     |
| ARTEMIS             | 5 Hz     | 0.25 Hz    |
| WIND                | 11 Hz    | 1 Hz       |
| STEREO              | 8 Hz     | 1 min      |

Since Juno does not provide direct plasma measurements during the cruise phase, we will estimate the spatial scale (thickness) of discontinuities using solar wind plasma state data derived from solar wind propagation models. Specifically, we will employ the Two-Dimensional Outer Heliosphere Solar Wind Modeling (MSWIM2D) [@keeblerMSWIM2DTwodimensionalOuter2022] to determine the ion bulk velocity ($v$) and plasma density ($n$) at Juno’s location. This model, which utilizes the BATSRUS MHD solver[@tothAdaptiveNumericalAlgorithms2012], simulates the propagation of solar wind from 1 to 75 astronomical units (AU) in the ecliptic plane, effectively covering the region pertinent to our study. The MSWIM2D model provides output data with an hourly time resolution as shown in @fig-model. The comparison of magnetic field magnitudes from MSWIM2D with those measured by Juno, after averaging to the same time resolution, reveals a strong correlation, confirming the model’s reliability.

::: {#fig-model}

![](figures/model/juno_model_validation_full)

**a,** Magnetic field magnitude from MSWIM2D and Juno. **b-c,** Plasma speed and density from MSWIM2D model. **d,** Juno radial distance from the Sun.
:::

## Methods {#sec-methods}

We employ @liuMagneticDiscontinuitiesSolar2022 method to identify discontinuities in the solar wind. This method provides better compatibility for discontinuities with small magnetic field changes and demonstrates greater robustness for conditions in the outer heliosphere compared to traditional methods, namely the $B$-criterion [@burlagaDirectionalDiscontinuitiesInterplanetary1969] and the $TS$-criterion [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. For each sampling time $t$, we define three intervals: the pre-interval $[-1,-1/2]\cdot T+t$, the middle interval $[-1/2,1/2]\cdot T+t$, and the post-interval $[1/2,1]\cdot T+t$, where $T$ represents time lags. The magnetic field time series in these three intervals are labeled as ${\mathbf B}_-$, ${\mathbf B}_0$, and ${\mathbf B}_+$, respectively. To identify a discontinuity, the following three conditions must be met:
$$
\begin{aligned}
\sigma({\mathbf B}_0) &> 2\max\left(\sigma({\mathbf B}_-), \sigma({\mathbf B}_+)\right) \\
\sigma\left({\mathbf B}_-+{\mathbf B}_+\right) &>\sigma({\mathbf B}_-)+\sigma({\mathbf B}_+) \\
|\Delta {\mathbf B}| &>|{\mathbf B}_{bg}|/10
\end{aligned}
$$
Here, $\sigma$ denotes the standard deviation, ${\mathbf B}_{bg}$ represents the background magnetic field magnitude, and $\Delta {\mathbf B}={\mathbf B}(t+T/2)-{\mathbf B}(t-T/2)$. The first two conditions ensure that the field changes are distinguishable from stochastic fluctuations, while the third condition serves as a supplementary measure to reduce recognition uncertainty [see discussion in @liuMagneticDiscontinuitiesSolar2022].

Notably, this method shares certain similarities with the Partial Variance of Increments (PVI) approach [@grecoPartialVarianceIncrements2017] for identifying coherent structures, as both calculate indices based on standard deviations (with this method focusing on the standard deviation of the magnetic field, and PVI on magnetic field increments). However, @liuMagneticDiscontinuitiesSolar2022 method is better suited for discontinuity identification by comparing data within neighboring time windows rather than across a correlation scale (an unknown quantity that may vary with radial distance and solar wind conditions). As a result, this approach effectively excludes wave-like coherent structures, eliminating the need for manual inspection and bypassing the necessity of grouping continuous clusters of high-PVI points as in PVI analysis.

Because the discontinuities may have different spatial scale at different radial distances from the Sun, we repeat the search procedure with time lags $T = 10, 20, 30, 40, 50, 60$ s, and combine the identified SWD events together to establish the event list.
Discontinuities with spatial scales larger than $60$ s are not considered in this study, however it could be argued that they are not significant in the outer heliosphere if they follow the trend with decreasing number of SWDs with increasing spatial scale [see @Vasko22:apjl; @Vasko24:jgr].
<!-- @fig-junoDistribution also indicates that the spatial scale of SWDs does not vary significantly with the radial distance from the Sun. -->

<!-- TODO: -->

More than ? SWDs are identified in the solar wind data from Juno, with ? SWDs identified within $[4, 5]$ AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal ($LMN$) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent function to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data (from model) to obtain the thickness and the current density of the discontinuity.
$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta t}\right) + c_l
$${#eq-fit}
where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity detection, $\Delta t$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

Previous studies have indicated that the minimum variance direction obtained through MVA is frequently unreliable \cite<see discussion in>{knetterFourpointDiscontinuityObservations2004,wangSolarWindCurrent2024}. Improved accuracy can be achieved by either using the cross product of the magnetic field measured at the boundaries as the normal direction [@knetterFourpointDiscontinuityObservations2004] or by enforcing conditions such as $Δ|B|/|B| > 0.05$ or $ω > 60°$ [@liuFailuresMinimumVariance2023], where $\Delta |{\mathbf B}|$ is the change in the magnitude of the magnetic field and $\omega$ is the field rotation angle across the discontinuity. The former approach requires a small normal component (component along the normal to SWD plane) in the magnetic field, while the latter approach limits the dataset to a subset that satisfies these conditions. In our analysis, we applied both methods and observed no significant differences in the results.

<!-- TODO: use both methods to determine the normal direction of the discontinuity and compare the results. -->

@fig-examples shows several examples of solar wind discontinuities detected by different spacecraft. Panels (a-c) show three examples of SWDs observed by Juno at 1, 3, and 5 AU distances, whereas panels (d-f) show three examples of SWDs observed by WIND, ARTEMIS, and STEREO-A (for these discontinuities we also show solar wind flow speed). All discontinuities share the same magnetic field configuration (in local $LMN$ coordinates): the main magnetic field component, $B_l$, reverses across the discontinuity, whereas the decrease of the magnetic pressure, $\sim B_l^2/2\mu_0$, is compensated by local enhancement of $B_m$ component. Such magnetic field rotation across the discontinuity keeps $|{\bf B}|$ almost constant \cite<typical magnetic field configuration for SWDs, see>{Artemyev19:grl:solarwind,vaskoKineticscaleCurrentSheets2022,Lotekar22,vaskoKineticScaleCurrentSheets2024}. The solar wind flow speed (three panels for WIND, ARTEMIS, and STEREO-A) shows typical jump across the discontinuities \cite<see discussion in>{Artemyev20:apjl}.

<!-- - We use [@liuMagneticDiscontinuitiesSolar2022] method to identify SWDs, which has better compatibility for the SWDs with minor field changes.
- Then the minimum variance analysis is applied to each ID event to obtain the boundary normal (LMN) coordinate and extract SWDs' features. -->

::: {#fig-examples}
![](figures/fig_examples)

Examples of solar wind discontinuities observed by various spacecraft at different heliocentric distances and times. Panels (a)–(c) display data from Juno at 1 AU (a), 3 AU (b), and 5 AU (c), showing magnetic field components $B_l$ (blue), $B_n$ (red), $B_m$ (green), and the magnetic field magnitude $B_{\text{total}}$ (black). Panels (d)–(f) show magnetic field and ion speed observations from other spacecraft: ARTEMIS on March 27, 2012 (d), STEREO on June 28, 2016 (e), and Wind on August 26, 2011 (f).
:::

# Results {#sec-results}

<!-- The basic approach of this proposal is to use solar wind measurements at 1AU (STEREO, ARTEMIS, WIND) to compare with Juno and PSP measurements and distinguish effects of solar wind temporal variations and effects of spatial (with the radial distance from the Sun) variations of discontinuity occurrence rate and characteristics. The example of such comparison for the occurrence rate is shown in @fig-rate, where we plot number of discontinuities measured per day by different spacecraft missions (for the same temporal resolution of magnetic field data and the same criteria of discontinuity determination). The radial distance of Juno for 2011-2016 is shown in @fig-model, and the number of discontinuities measured by Juno per day coincides with the discontinuity number measured by STEREO, WIND, and ARTEMIS, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance (with time after $\sim 2013$), as Juno moves from $1$ AU to $5$ AU. We will use the similar comparison for discontinuity characteristics and occurrence rate derived for PSP and Juno.  -->

## Waiting time

To comprehensively understand the radial evolution of the SWD occurrence rate, particularly the average occurrence rate, it is essential to determine the statistical distributions underlying the observed samples. This enables the derivation of meaningful inferences. In this context, we derived the probability distribution of the waiting time $τ = t_{0,i} - t_{0,i-1}$, where $t_{0,i}$ is the time of detection of the $i$-th discontinuity event at various radial distances ($τ$ is also referred to as the interarrival interval in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979).

To model the data, we employ the Weibull distribution [@weibullStatisticalDistributionFunction1951] due to its flexibility in capturing varying occurrence rates. This choice allows for a more accurate representation and an improved fit compared to the exponential distribution previously used in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979. The probability density function of the Weibull distribution is given as follows:
$$
f(τ; α, θ) = \frac{α}{θ} \left( \frac{τ}{θ} \right)^{α-1} e^{-(τ/θ)^α}, \quad τ \ge 0
$$
where $α$ and $θ$ represent the shape and scale parameters, respectively. The complementary cumulative distribution function of the Weibull distribution $F(τ; α, θ) = 1 - \mathrm{e}^{-(τ/θ)^α}$ is a stretched exponential function (The Weibull distribution reduces to the exponential distribution when $α = 1$).

::: {#fig-waitingTime}
![](figures/fig_wt_dist.pdf)

Waiting time probability density functions $p(\tau)$ for Juno at 1 AU in 2011 (top) and 5 AU in 2016 (bottom). Observed data (black) are fitted with Weibull (blue) and exponential (orange) distributions. In 2011, the Weibull distribution is characterized by parameters $\alpha = 0.99, \theta = 30.29$, and a mean waiting time of 30.41 minutes, while the exponential distribution has a mean of 30.42 minutes. In 2016, the Weibull distribution has parameters $\alpha = 0.65, \theta = 110.50$, and a mean of 150.96 minutes, whereas the exponential distribution has a mean of 165.02 minutes. Vertical dashed lines denote the mean waiting times for each fitted distribution.
:::

The observed waiting time distribution, along with the fitted Weibull distribution, is shown in @fig-waitingTime. For $α < 1$, the likelihood of an event occurring decreases with increasing waiting time. This behavior aligns with our expectations for the discontinuity occurrence rate in response to the evolving state of the solar wind. Across the largest range of waiting times, the observed probabilities exceed those predicted by the fitted distribution. This discrepancy may be attributed to data gaps, which artificially extend the observed waiting times between discontinuities.

The mean waiting time $\bar{τ}$ can be determined by integrating the probability density function $\bar{τ} = \int_0^\infty τ f(τ) dτ$. The mean occurrence rate, in turn, is simply the reciprocal of the mean waiting time. Notably, the mean value derived from the fitted Weibull distribution is more reliable, as it mitigates the influence of unusually large waiting times caused by data gaps.

## Occurrence rate

The occurrence rate of SWDs is investigated with regard to the radial distance from the Sun for Juno, ARTEMIS, STEREO, and Wind. The number of discontinuities measured per day by different spacecraft missions (for the same temporal resolution of the magnetic field data and the same criteria of discontinuity determination) is shown in @fig-rate. Because data is not continuously available and ARTEMIS spacecrafts will only be exposed in the solar wind for a relative short time during its orbits around the Moon (the Moon would be in the magnetosphere for a $\sim 30$% of it orbital time), the occurrence rate (number of SWDs per day) is corrected with respect to data gap and the time the mission spent in the pristine solar wind. 
The occurrence rate of SWDs observed by Juno is consistent with the occurrence rate observed by 1AU missions, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance as Juno moves from $1$ AU to $5$ AU (corresponding to time after $\sim 2013$, the radial distance of Juno for 2011-2016 is shown in @fig-model)

The rate of occurrence depends linearly on the solar wind velocity [@sodingRadialLatitudinalDependencies2001], however the plasma velocity data in Juno mission is not available during its cruise phase. Therefore, we do not normalize the occurrence rate with respect to the solar wind speed. This influence of the solar wind speed on the occurrence rate might be mitigated by the fact all the spacecrafts are in the ecliptic plane, and the solar wind speed is not expected to change significantly with the radial distance from the Sun beyond $1$ AU. We normalize the occurrence rate from Juno to the occurrence rate from 1AU missions to account for the effect of solar wind structures on the occurrence rate. 
The normalized occurrence rate is shown in panel (b) of @fig-rate. The occurrence rate of SWDs drops with the radial distance from the Sun, following $1/r$ law. So the decrease of the occurrence rate with increasing radial distance from the Sun is expected to be a real effect of the radial distance from the Sun and this confirms the former results [@sodingRadialLatitudinalDependencies2001].

::: {#fig-rate layout-nrow="1"}

![](figures/ocr/ocr_time_cleaned)

![](figures/ocr/ocr_r_N2_cleaned_simplified.pdf)

The number of discontinuities measured by Juno per day coincides with the discontinuity number measured by STEREO, WIND, and ARTEMIS, when Juno is around $1$ AU. This number (occurrence rate) decreases with distance (with time after $\sim 2013$), as Juno moves from $1$ AU to $5$ AU. The radial distance of Juno for 2011-2016 is shown in @fig-model.
:::

## Current density and thickness

The current density $J_m$ supporting the main magnetic field reversal across a discontinuity is determined by 
$$
J_m = - \frac{1}{μ_0 V_n} \frac{d B_l}{d t},
$$ {#eq-j}
where $μ_0$ is the vacuum permeability and $V_n$ is the normal component of the proton flow velocity. For fitting method, the derivative of the magnetic field could be estimated as $\max(d B_l / d t) = B_{l0} / \Delta t$. This estimate is more robust than the direct calculation of the derivative of the magnetic field, because the fitting method is less sensitive to the noise in the data and the time resolution of the data (see discussion in the Appendix).

<!--  The proton flow velocity $v$ is expressed as $v = v_{sw} - v_{s/c} + U$, where $v_{sw}$ is the solar wind velocity, $v_{s/c}$ is the spacecraft velocity, and $U$  denotes the propagation speed of the discontinuities. In our case, the latter two velocities are negligible. -->

The thickness $d$ of a discontinuity is determined by

$$
d = V_n \Delta t
$$

where $\Delta t$ is the transition time derived from the fitting method.

To understand the relationship between the discontinuity and the local plasma, we normalize the thickness of the discontinuity by the ion (proton) inertial length $d_i=c/w_{pi}$ and the current density by the Alfven current density $J_A = e N_p V_A$ (corresponding to the drift between protons and electrons of local Alfén speed $V_A$). @fig-junoDistribution presents the statistical distribution of the (un)normalized thickness and current density of SWDs grouped by the radial distance from the Sun collected by Juno. 
Panel (a,b) shows that the thickness increases with radial distance. However, after normalization, the thickness slightly decreases, with the most probable value around 4-2 $d_i$. With increasing radial distance, larger fraction of SWDs have thickness less than 1 $d_i$, becoming sub-proton scale structures.  Panel (c,d) shows that the current density decreases with increasing radial distance. However, after normalization, it slightly increases, with the most probable value around 0.05-0.15 $J_A$. It should be noted that for the normalized thickness and current density, the change  with the radial distance is not significant compared to the spread of the distribution [see also discussion in @vaskoKineticScaleCurrentSheets2024].

The distribution of the thickness and current density of SWDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) during the five year of Juno cruise phase are shown in @fig-windDistribution. The distribution are grouped by the year of observation. Panels (a,b) show that both the thickness and normalized thickness distribution of SWDs remains almost constant with the year of observation. Most of the discontinuities have thickness around 350-3500 km at $1$ AU, with the most probable value around 1000 km. Panels (c,d) show that the current density slightly increases with the year of observation, with the most probable value around 5 $nA m^{-2}$, in consistence with previous studies [@vaskoKineticscaleCurrentSheets2022]. However, after normalization, the current density distribution remains identical across the five years. The change in the current density could be explained by the change in the solar wind properties with the solar cycle, as shown in @fig-swParameters. Density and magnetic field associated with the discontinuities increases with the solar cycle approaching the solar minimum, which could lead to the increase of the current density of the discontinuities with fixed normalized current density.

::: {#fig-junoDistribution}
![](figures/juno_distribution_r)

Distribution of various properties of SWDs observed by Juno grouped by the radial distance from the Sun. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

::: {#fig-windDistribution}
![](figures/wind_distribution_time)

Distribution of various properties of SWDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) grouped by the year of observation. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

::: {#fig-swParameters}
![](figures/wind_sw_paramters)

Solar wind parameters associated with the SWDs observed by 1AU satellites (Wind, ARTEMIS and STEREO-A) grouped by the year of observation. Panel (a) solar wind density, (b) magnetic field, (c) fitted magnetic field $A$ in @eq-fit.
:::

# Discussion

## Spatial variations of discontinuity properties

TODO:

## Dependence of occurrence rate {#sec-dependence}

The observed increase in the average waiting time between discontinuities with radial distance prompts us to explore different hypotheses to explain this phenomenon. At any given radial distance, discontinuities may originate from two primary sources: those generated near the Sun and convected outward by the solar wind, and those formed locally. Since the solar wind speed is significantly higher than the Alfvén speed—which characterizes the propagation speed of rotational discontinuities—the fraction of inward-propagating discontinuities is expected to be negligible. Additionally, when observations are made by a single spacecraft, some discontinuities may be missed due to unfavorable positioning or limitations in detection methods.

First, we consider the limitations of our current detection methods. An increase in the duration of discontinuities could render some of them unidentifiable if their duration becomes comparable to the time interval used in our analysis, $T$. Although we observe a slight increase in the average duration of discontinuities with radial distance, this increase is relatively small. The overall population of discontinuities is still predominantly characterized by those with short durations (see @fig-TEffect in the Appendix). Since our method normalizes the change in the magnetic field by the surrounding magnetic field, a decrease in magnetic field variation with radial distance alone does not necessitate non-identification. Notably, our results differ from previous studies [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979], which reported typical durations of 2 to 3 seconds at 1 AU and 10 to 30 seconds at 5 AU. In our analysis, however, the observed durations remain below 10 seconds at 5 AU, suggesting that our method may be more effective at identifying shorter-duration events than those reported before.

If the majority of distant discontinuities are generated locally, and assuming that the primary conditions for their generation do not change significantly with radial distance, we would expect them to be uniformly distributed throughout the hemisphere. Consequently, the occurrence rate should be independent of radial distance. To test this hypothesis, we examined the relationship between waiting time and plasma beta. However, even at a 99% statistical significance level, we could not reject the null hypothesis that these two variables are independent. Therefore, if discontinuities are primarily generated locally, there must be other factors influencing their generation rates.

Alternatively, we consider the hypothesis that the majority of discontinuities are generated close to the Sun. A realistic assumption is that discontinuities are three-dimensional curved surfaces with lengths and widths proportional to their thickness. In this scenario, the occurrence rate observed by a single spacecraft does not represent the total number of discontinuities within a given shell (radial distance). Since spacecraft travels within the ecliptic plane and the normal direction of most discontinuities lies within this plane, it is necessary to project the discontinuities onto a circular cross-section. The projected lengths should then be normalized by the circumference of the circle to estimate the total number of discontinuities within the cylindrical shell surrounded Sun at given radial distance. The normalization factor is  $d \sin\theta_{n,v} / r$ , where $d$ is the thickness of the discontinuity, $\theta_{n,v}$ is the angle between the normal direction of the discontinuity and the solar wind velocity, and $r$ is the radial distance from the Sun (the result is shown in the Appendix). Accounting for this factor reduces the differences in the normalized occurrence rates observed by different spacecraft. However, we cannot rule out the possibility that some discontinuities may be annihilated as they convect outward with the solar wind. This annihilation process would result in a lower number of discontinuities observed by Juno at larger radial distances. Therefore, under this hypothesis, the annihilation process must play a significant role. (However, if we disregard changes in the thickness and only account for the factor of the circumference of the circle, the normalized discontinuities observed by Juno would be consistent with those observed at 1 AU.)

![](figures/ocr/ocr_normalized_by_r)

# Conclusion

We have collected 5 years of solar wind discontinuities from Juno, Wind, ARTEMIS and STEREO. The results of analysis of these datasets can be summarized as follows:

1. The normalized occurrence rate of SWDs drops with the radial distance from the Sun, following $1/r$ law within $[1,5]$AU.
2. The thickness of SWDs increases with the radial distance from the Sun, but after normalization to ion inertial length, the thickness of SWDs slightly decreases or remains almost constant within the scale of thickness distribution. The thickness distribution of SWDs remains almost constant with the time scale of one year of observation.
3. The current density of SWDs decrease with the radial distance from the Sun, but after normalization to the Alfven current, the current density of SWDs slightly increases or remains almost constant within the scale of thickness distribution. The current density distribution of SWDs slightly increases with the year of observation from 2011 to 2016, but after normalization, the current density distribution remains identical across the five years, thus no effect of the long-term (i.e. solar cycle) temporal variations is detected for normalized current density.

# Acknowledgements

Sunspot data from the World Data Center SILSO, Royal Observatory of Belgium, Brussels. ARTEMIS solar wind time interval data from the OMNI database. ARTEMIS solar wind time interval data from OMNIWEB, link: [omniweb.gsfc.nasa.gov/ftpbrowser/themis_b_sw.txt](https://omniweb.gsfc.nasa.gov/ftpbrowser/themis_b_sw.txt).

# Open Research

Codes and datasets used for the analysis are available in the GitHub repository https://github.com/Beforerr/ids_spatial_evolution_juno.

# Appendix

## Effect of the time resolution on the discontinuity properties

The transition time $\Delta t$ is determined in one second resolution data for all our missions to reduce the system error. (Note that highest time resolution of the data is not needed for the transition time determination by fitting method).

![Effect of the time resolution on the discontinuity properties](figures/ts_effect){#fig-tsEffect}

![The relative percentage of events as a function of radial distance (in AU) for different values of the parameter $T$ (20s, 30s, 40s, 50s, and 60s). The relative percentage of events is normalized to the number of events using $T=20$s.](figures/ocr/juno_tau_r.pdf){#fig-TEffect}