### 2025-03-13 notebooks/102_results_scattering.qmd
update scattering

---
title: Ion Scattering by Solar Wind Discontinuities
---

```{julia}
include("../src/main.jl")
using Dates
using Beforerr
using GLMakie
using FHist
using AlgebraOfGraphics: aog_theme

M = var_mapping()
```

Loading the data

```{julia}
name = "Wind"
df = load(; name);
```

```{julia}
# velocity v is normalized to $v0 = l Ω$, where $l$ represents the width of the RD and $Ω = q B/(mpc)$ is the cyclotron frequency.
using Unitful: ε0, μ0, mp, q, c
q_G = q / sqrt(4π * ε0)
B_G_factor = sqrt(4π / μ0)

function transform_scattering!(df::AbstractDataFrame)
    # gyrofrequency
    @chain df begin
        @transform!(
            :θ = acosd.(:bn_over_b),
            :E_norm = @. (q_G * :L_n_cross * :"B.mean" * B_G_factor / (mp * c^2)) |> NoUnits
        )
    end
end

transform_scattering!(df);
```

```{julia}
binedges = (-6:1, 0:10:180, 0:10:180)
hist3dbin(df) = Hist3D(
    (log10.(df[!, :E_norm]), df[!, :ω_in], df[!, :θ]); binedges
)

to_endpoints(binedges) = (first(binedges), last(binedges))

let
    h3d1 = hist3dbin(subset(df, :accuracy))
    h3d2 = hist3dbin(subset(df, :accuracy => ByRow(!)))
    z = zeros(eltype(bincounts(h3d1)))
    # axis = (; aspect=:equal, xlabel=L"\log (ṽ_B)", ylabel="In-plane rotation angle", zlabel="Azimuthal angle")
    axis = (; aspect=:equal, xlabel=L"\log (ṽ_B)", ylabel=L"ω_{in}", zlabel=L"θ")

    f = Figure(; size=(600, 300), figure_padding=0)
    ax1 = Axis3(f[1, 1]; axis..., title="Reliable", protrusions=(30, 0, 60, 30))
    volume!(ax1, to_endpoints.(h3d1.binedges)..., replace(bincounts(h3d1), z => NaN))
    # volume!(ax1, h3d1)
    ax2 = Axis3(f[1, 2]; axis..., title="Unreliable", protrusions=0)
    p = volume!(ax2, to_endpoints.(h3d2.binedges)..., replace(bincounts(h3d2), z => NaN))
    Colorbar(f[1, 3], p)
    # volume!(ax2, h3d2)
    f
end

easy_save("scatter/$(name)_hist3d"; formats=[:png])
```

![](../figures/scatter/wind_hist3d.png)

```{julia}
# Volume plot type support in AOG
# https://github.com/MakieOrg/AlgebraOfGraphics.jl/pull/572

θ = :θ => "azimuthal angle"
Bn = :bn_over_b => abs => L"B_N/B";
v0 = :v0 => log10 => "Log v0";

layer = data(df) * mapping(layout=:accuracy);

xyz = (:E_norm, M.ω_in, θ)
datalimits = ((1, 14), (0, 180), (0, 180))
axis = (type=Axis3, width=400)
v = histogram(; datalimits, normalization=:pdf)
plt = data(df) * mapping(xyz...; layout=:accuracy) * v
fg = draw(plt; axis)
colorbar!(fg)
fg
```

```{julia}
datalimits = (2, 14)
axis = (; yscale=log10)
v = histogram(; normalization=:pdf, bins=20, datalimits)
dfplot(layer * v, v0; axis)
```

```{julia}
using FHist
using LinearAlgebra

function get_data(df; x=:v0, y=:ω_in, z=:θ)
    x = df[!, x] .|> ustrip .|> log2
    y = df[!, y]
    z = df[!, z] .|> abs
    return x, y, z
end

function get_hist3d(df)
    return Hist3D(
        get_data(df);
        binedges=(1.5:14.5, 20:10:180, 0:10:90),
    ) |> normalize
end

function plot_hist3d(df)
    h = get_hist3d(df)
    fig = Figure()
    ax = Axis3(fig[1, 1]; width=400, xlabel="log2(v0)", ylabel="ω_in", zlabel="θ")
    plot!(h)
    fig
end

using GLMakie
GLMakie.activate!()
plot_hist3d(filter_low_accuracy(df))
```

## Export

```{julia}
using HDF5

filename = datadir("$(name)_hist3d.h5")
h = get_hist3d(filter_low_accuracy(df))

function save(filename, h::Hist3D)
    # delete file if it exists
    isfile(filename) && rm(filename)
    h5writehist(filename, "hist", h)
end

save(filename, h)
```

```{julia}
*(size(bincounts(h))...) |> println
bincenters(h)
```