### 2025-04-25 notebooks/02_waiting_time.qmd
docs: update

---
title: Waiting time analysis
---

[Archived notebook using Python and R](analysis/01_occurence_rate.ipynb)

## Notes

- Occurence rate is related to the waiting time distribution.
- Group the data by the radial distance for WT analysis is not good as the radial distance is not a monotonic function of time.

It is hard to remove duplicate events when combining multiple datasets with different τ. So when using the combined dataset, the occurence rate would be overestimated.

> An analysiswas carried out to determine the probability distribution governing the time interval between successive discontinuities. The times at which the discontinuities occurred,obtainedfromthe identificationprogram, were used to computethe time difference, $τ = T_j - T_{j-1}$, the so-called interarrival interval. The number of cases were then tabulated corresponding to discreteranges of τ, and a histogramwas prepared.By properly normalizing the numberof casesin each range a probability distribution function giving the relative frequency of occurrencewas obtained. A similar study was previously carried out by Burlaga[1969].

### References

- [Weibull distribution - Distributions.jl](https://juliastats.org/Distributions.jl/stable/univariate/#Distributions.Weibull)

$$
f(τ; α, θ) = \frac{α}{θ} \left( \frac{τ}{θ} \right)^{α-1} e^{-(τ/θ)^α}, \quad τ \ge 0
$$

```{julia}
using DrWatson
@quickactivate
include(srcdir("main.jl"))

using Discontinuity: remove_duplicates
using CairoMakie
```

```{julia}
df = load_taus(; unique_f=missing)

wind_df = load(; name="Wind", ts=missing);
jno_df = load();
```

```{julia}
subset_jno(df, r) = subset(df, :radial_distance => x -> round_c.(x) .== r)
subset_time(df, t) = subset(df, :time => x -> x .< t)
```

```{julia}
df = wind_df;
df_s1 = subset(df, :time => t -> year.(t) .== 2013);
df_s2 = subset(df_s1, :time => t -> month.(t) .< 2);

jno_df_s1 = subset(jno_df, :time => t -> year.(t) .== 2011);
jno_df_s2 = subset(jno_df, :time => t -> year.(t) .== 2016);
```

## Waiting time distribution

```{julia}
using Distributions

function plot_wt_pdf_lim(df; xlims=(1, 1000), ylims=(1e-5, 8e-2), kwargs...)
    f = plot_wt_pdf(df; kwargs...)
    xlims!(xlims...)
    ylims!(ylims...)
    return f
end

using Printf
Base.show(io::IO, f::Float64) = @printf(io, "%.2f", f)

dists = (Weibull, Exponential)
step = 5
```

```{julia}
#| column: page
#| layout-ncol: 3
dfs = [jno_df, jno_df_s1, jno_df_s2]
fgs = plot_wt_pdf_lim.(dfs; dists)
display.(fgs);
```

```{julia}
using Beforerr: DistsFit

WaitingTimeDist(df, args...) = DistsFit(waiting_time(df), args...)
```

```{julia}
function plot_fig_wt_dist(; step=5, df1=jno_df_s1, df2=jno_df_s2)
    limits = (0, 680, 5e-6, 5e-2)
    f = Figure()
    new_ax(pos; kw...) = Axis(f[pos...]; yscale=log10, xlabel="τ (minutes)", ylabel="p(τ)", limits, kw...)

    ax1 = new_ax((1, 1); xlabel="")
    plot!(WaitingTimeDist(df1, dists, step))
    ax2 = new_ax((2, 1);)
    plot!(WaitingTimeDist(df2, dists, step))
    axislegend(ax1, "Juno, 2011"; loc=:upperright)
    axislegend(ax2, "Juno, 2016"; loc=:upperright)
    return f
end

# let 
#     plot_fig_wt_dist()
#     easy_save("fig_wt_dist")
# end

let tdf = Discontinuity.remove_duplicates(df)
    df1 = subset(tdf, :time => t -> year.(t) .== 2011)
    df2 = subset(tdf, :time => t -> year.(t) .== 2016)
    plot_fig_wt_dist(; df1, df2)
    easy_save("fig_wt_dist_no_duplicates")
end
```

## Occurence rate

```{julia}
using Distributions
import DataFrames: groupby

get_fit_params(x; dist=Weibull, func=mean) = func(fit(dist, x))
get_wt_fit_params(time; kwargs...) = get_fit_params(waiting_time(time); kwargs...)

function groupby(df::AbstractDataFrame, dt::Period; col=:time)
    @chain df begin
        @transform!(:time_bin = round.(:time, dt),)
        groupby(:time_bin)
    end
end

begin
    dt = Day(50)
    tdf = @chain df begin
        remove_duplicates
        groupby(dt)
        combine(
            nrow,
            :time => get_wt_fit_params => :τ,
            [:r] .=> mean
        )
        @transform(:ρ = 1 ./ :τ)
    end

    x = :time_bin => "Time"
    specs = data(tdf) * mapping(x, :ρ => "Occurence rates")
    draw(specs)
end
```

### Normalization by thickness

```{julia}
time_interval = Day(160)

jno_ρ_df = @chain jno_df begin
    @transform(
        :time_bin = (:time - minimum(:time)) ./ time_interval .|> round,
        :l = :L_k .* (abs ∘ cosd).(:θ_vn),
        :l_n = :L_k .* (abs ∘ cos ∘ atan).(:Vn_y ./ :Vn_x)
    )
    @transform(
        :l_ratio = :l ./ :r,
        :l_n_ratio = :l_n ./ :r
    )
    groupby(:time_bin)
    combine(
        nrow,
        :time => get_wt_fit_params => :τ,
        [:r :l :l_ratio :l_n_ratio] .=> mean
    )
    # subset(:nrow => x -> x .> 2000)
    @transform(
        :ρ = 1 ./ :τ,
    )
    @transform(
        :ρ_l = :ρ ./ :l_ratio_mean * maximum(:l_ratio_mean),
    )
    sort(:r_mean)
end
```

```{julia}
wind_ρ_df = @chain wind_df begin
    @transform!(
        :time_bin = (:time - minimum(:time)) ./ time_interval .|> round,
    )
    groupby(:time_bin)
    combine(
        nrow,
        :time => get_wt_fit_params => :τ,
    )
    @transform!(
        :ρ = 1 ./ :τ,
    )
end
```

```{julia}
df = leftjoin(jno_ρ_df, wind_ρ_df, on=:time_bin, makeunique=true)
df.ρ_ratio = @. df.ρ / df.ρ_1

@transform!(df,
    :ratio = :ρ_ratio ./ :l_ratio_mean,
    :ratio_n = :ρ_ratio ./ :l_n_ratio_mean,
)
@transform!(
    df,
    :ratio = :ratio ./ mean(:ratio),
    :ratio_n = :ratio_n ./ mean(:ratio_n)
)

df
```

```{julia}
x = :time_bin => "Time"
legends = ["Juno", "Wind"]
specs = data(df) * mapping(x, [:ρ_l, :ρ_1] .=> "Normalized Occurence rates") * mapping(color=dims(1) => renamer(legends))
draw(specs)
easy_save("ocr/ocr_normalized_by_r")
```

```{julia}
x = :r_mean => r_lab
specs = data(df) * mapping(x, :ρ)
draw(specs)
```

```{julia}
y = [:ratio_n, :ratio]
specs = data(df) * mapping(x) * mapping(y, color=dims(1) => renamer(string.(y)))
plt2 = draw(specs)
```

## Compare the exponential and Weibull fits for the waiting time distributions of Wind dataset

```{julia}
f1 = plot_wt_pdf_lim(wind_df, ylims=(1e-6, 1e-1),)
f2 = plot_wt_pdf_lim(wind_df, dist=Weibull, ylims=(1e-6, 1e-1))
display.([f1, f2]);
```

```{julia}
df1 = subset_time(jno_df, Date(2011, 9, 30))
df2 = subset_time(wind_df, Date(2011, 9, 30))
dfs = [df1, df2]
fgs = dfs .|> plot_wt_pdf_lim
display.(fgs);
```