### 2025-04-25 notebooks/21_model.qmd

# MHD Model and Juno solar wind parameters comparison

@wilsonSolarWindProperties2018

```{julia}
using DrWatson
using Discontinuity
using Discontinuity: Alfven_current
using DataFramesMeta
using PlasmaFormulary: inertial_length
using SpaceDataModel: tryparse_datetime

@quickactivate
include("../src/main.jl")
```

```{julia}
using CSV
using DataFrames
using Unitful

"""
    read_sw_params_csv(file="jgra54158-sup-0001-supinfo.csv")

Reads the solar wind parameters CSV file and returns a DataFrame.
"""
function read_sw_params_csv(file=datadir("jgra54158-sup-0001-supinfo.csv"))
    df = CSV.read(file, DataFrame)
    @chain df begin
        @transform! @byrow begin
            :time = tryparse_datetime(String(:UTC))
            :N_PROTONS_CC = :N_PROTONS_CC * u"cm^-3"
            :V_KMPS = :V_KMPS * u"km/s"
        end
    end
    return df
end

@b read_sw_params_csv()

# jno_df = load()
```

```{julia}
using FlexiJoins
using Distances
using Dates
using Unitful: μ0

jno_df_with_params = innerjoin(
    (jno_df, jno_sw_params),
    by_pred(x -> x.time .. (x.time + Minute(1)), ∋, :UTC);
)

@chain jno_df_with_params begin
    @transform! begin

        :d_star = :d_star .* u"nT/s"
        :duration = :duration .* u"s"
        :V_KMPS = :V_KMPS .* u"km/s"
    end
    @transform! begin
        :V_n_cross_sw = @. abs(first(:n_cross) * :V_KMPS)
        :d_i_sw = inertial_length.(:N_PROTONS_CC, Unitful.q, Unitful.mp)
        :j_Alfven_sw = Alfven_current.(:b_mag, :N_PROTONS_CC)
    end
    @transform! begin
        :L_n_cross_sw = :V_n_cross_sw .* :duration
        :j0_k_sw = @. abs(:d_star / :V_n_cross_sw / μ0) |> u"nA/m^2"
    end
    @transform! begin
        :L_n_cross_sw_norm = @. :L_n_cross_sw / :d_i_sw |> NoUnits
        :j0_k_sw_norm = @. :j0_k_sw / :j_Alfven_sw |> NoUnits
    end
end
```

```{julia}
using CairoMakie
using UnitfulLatexify: latexroundunitlabel as lru

let df = jno_df_with_params, xlabel = "Model", ylabel = "Juno"
    f = Figure()
    ax = Axis(f[1, 1]; title=lru("V", u"km/s"), ylabel)
    scatter!(ax, norm.(df[!, :V]), ustrip.(df[!, :V_KMPS]))

    ax2 = f[1, 2] = Axis(f; title=lru("n", u"cm^-3"), xscale=log10, yscale=log10)
    scatter!(ax2, ustrip.(df[!, :n]), ustrip.(df[!, :N_PROTONS_CC]))
    ax3 = Axis(f[2, 1]; title=lru("L", u"km"), ylabel, xscale=log10, yscale=log10)
    scatter!(ax3, ustrip.(df[!, :L_n_cross]), ustrip.(df[!, :L_n_cross_sw]))
    ax4 = f[2, 2] = Axis(f; title=lru("J", u"nA/m^2"), xscale=log10, yscale=log10)
    scatter!(ax4, ustrip.(df[!, :j0_k]), ustrip.(df[!, :j0_k_sw]))
    ax5 = f[3, 1] = Axis(f; title=L"L/d_i", xlabel, ylabel, xscale=log10, yscale=log10)
    scatter!(ax5, df[!, :L_n_cross_norm], df[!, :L_n_cross_sw_norm])
    ax6 = f[3, 2] = Axis(f; title=L"J/J_{Alfven}", xlabel, xscale=log10, yscale=log10)
    scatter!(ax6, df[!, :j0_k_norm], df[!, :j0_k_sw_norm])
    for ax in (ax,)
        ablines!(ax, 0, 1, color=:red, linestyle=:dash)
    end

    lines!(ax2, exp10.(-1:0), exp10.(-1:0), color=:red, linestyle=:dash)
    lines!(ax3, exp10.([2, 4.5]), exp10.([2, 4.5]), color=:red, linestyle=:dash)
    lines!(ax4, exp10.(-2:1), exp10.(-2:1), color=:red, linestyle=:dash)
    lines!(ax5, exp10.(-1:2), exp10.(-1:2), color=:red, linestyle=:dash)
    lines!(ax6, exp10.(-2:1), exp10.(-2:1), color=:red, linestyle=:dash)

    easy_save("model/juno_sw_comparision")
end
```

![](../figures/model/juno_sw_comparision.png)