### 2025-04-30 notebooks/30_data.qmd
feat: load Juno data

---
title: Data Processing (Julia)
---

```{julia}
using Parquet2
using DrWatson
using DataFrames
using DimensionalData

function get_juno_data(; dir=datadir(), cols=["BX SE", "BY SE", "BZ SE"])
    ds = Parquet2.Dataset(dir * "/2011.parquet")
    data = reduce(hcat, Parquet2.load(ds, col) for col in cols)
    times = Parquet2.load(ds, "time")
    dims = (Ti(times), Y())
    return DimArray(data, dims)
end

b = get_juno_data()
```

```{julia}
using Discontinuity
using Dates
period = Second(60)

# this is to speed up the detection
# also Juno data is of low precision
_float16(x) = convert(Array{Float16,ndims(x)}, x)

detect_variance(_float16(b), t, period, 10, Val(1))
```