### 2025-05-03 notebooks/21_model.qmd
update

# MHD Model and Juno solar wind parameters comparison

@wilsonSolarWindProperties2018

```{julia}
using Discontinuity
using Discontinuity: compute_params!, Alfven_current, inertial_length, gradient_current
using Juno: produce_or_load_juno
using PlasmaFormulary: CurrentDensity, NumberDensity
```

```{julia}
using DataFrames, DataFramesMeta
using Unitful
using Juno.Data: read_sw_params_csv

jno_df = produce_or_load_juno()
jno_sw_params = read_sw_params_csv(; add_unit=true)
```

```{julia}
using FlexiJoins
using Dates
using Unitful: μ0

jno_df_with_params = innerjoin(
    (jno_df, jno_sw_params),
    by_pred(x -> x.time .. (x.time + Minute(1)), ∋, :time);
)

@chain jno_df_with_params begin
    compute_params!
    @transform! begin
        :V_n_cross_sw = @. abs(first(:n_cross) * :V_KMPS)
        :V_n_mva_sw = @. abs(first(:n_mva) * :V_KMPS)
        :d_i_sw = inertial_length.(:N_PROTONS_CC, Unitful.q, Unitful.mp)
        :J_A_sw = Alfven_current.(:B_mag, :N_PROTONS_CC)
    end
    @transform! @astable begin
        :L_n_cross_sw = :V_n_cross_sw .* :duration
        :L_n_cross_sw_norm = @. :L_n_cross_sw / :d_i_sw |> NoUnits
        :L_n_mva_sw = :V_n_mva_sw .* :duration
        :L_n_mva_sw_norm = @. :L_n_mva_sw / :d_i_sw |> NoUnits
        :J_m_max_cross_sw = @. abs(gradient_current(:grad, :V_n_cross_sw))
        :J_m_max_cross_sw_norm = @. :J_m_max_cross_sw / :J_A_sw |> NoUnits
        :J_m_max_mva_sw = @. abs(gradient_current(:grad, :V_n_mva_sw))
        :J_m_max_mva_sw_norm = @. :J_m_max_mva_sw / :J_A_sw |> NoUnits
    end
end
```

```{julia}
using CairoMakie
using UnitfulLatexify: latexroundunitlabel as lru
using CairoMakie.GeometryBasics

p_big = decompose(Point2f, Circle(Point2f(0), 1))
p_small = decompose(Point2f, Circle(Point2f(0), 0.5))

uless(x) = NoUnits(x)
uless(x::NumberDensity) = NoUnits(x / u"cm^-3")
uless(x::CurrentDensity) = NoUnits(x / u"nA/m^2")
uless(x::Unitful.Length) = NoUnits(x / u"km")
uless(x::Unitful.Velocity) = NoUnits(x / u"km/s")

let df = jno_df_with_params, xlabel = "Model", ylabel = "Juno", xscale = log10, yscale = log10, plot = (; alpha=0.3)
    f = Figure(; size=(600, 600))
    ax1 = Axis(f[1, 1]; title=lru("V", u"km/s"), ylabel)
    ax2 = Axis(f[1, 2]; title=lru("n", u"cm^-3"), xscale, yscale)
    ax3 = Axis(f[2, 1]; title=lru("L", u"km"), ylabel, xscale, yscale)
    ax4 = Axis(f[2, 2]; title=lru("J", u"nA/m^2"), xscale, yscale)
    ax5 = Axis(f[3, 1]; title=L"L/d_i", xlabel, ylabel, xscale, yscale)
    ax6 = Axis(f[3, 2]; title=L"J/J_{Alfven}", xlabel, xscale, yscale)

    p1, p2 = @with df begin
        scatter!(ax1, uless.(:V_mag), uless.(:V_KMPS))
        scatter!(ax2, uless.(:n), uless.(:N_PROTONS_CC))

        p1 = scatter!(ax3, uless.(:L_n_cross), uless.(:L_n_cross_sw), label="Cross"; plot...)
        p2 = scatter!(ax3, uless.(:L_n_mva), uless.(:L_n_mva_sw), label="MVA"; plot...)

        scatter!(ax4, uless.(:J_m_max_cross), uless.(:J_m_max_cross_sw), label="Cross"; plot...)
        scatter!(ax4, uless.(:J_m_max_mva), uless.(:J_m_max_mva_sw), label="MVA"; plot...)

        scatter!(ax5, uless.(:L_n_cross_norm), uless.(:L_n_cross_sw_norm); plot...)
        scatter!(ax5, uless.(:L_n_mva_norm), uless.(:L_n_mva_sw_norm); plot...)
        scatter!(ax6, uless.(:J_m_max_cross_norm), uless.(:J_m_max_cross_sw_norm); plot...)
        scatter!(ax6, uless.(:J_m_max_mva_norm), uless.(:J_m_max_mva_sw_norm); plot...)
        (p1, p2)
    end

    Legend(f[1, 3], [p1, p2], ["Cross", "MVA"])

    for ax in (ax1,)
        ablines!(ax, 0, 1, color=:red, linestyle=:dash)
    end

    lines!(ax2, exp10.(-1:0), exp10.(-1:0), color=:red, linestyle=:dash)
    lines!(ax3, exp10.([2, 4.5]), exp10.([2, 4.5]), color=:red, linestyle=:dash)
    lines!(ax4, exp10.(-2:1), exp10.(-2:1), color=:red, linestyle=:dash)
    lines!(ax5, exp10.(-1:2), exp10.(-1:2), color=:red, linestyle=:dash)
    lines!(ax6, exp10.(-2:1), exp10.(-2:1), color=:red, linestyle=:dash)

    limits!(ax1, 360, 550, 360, 550)

    easy_save("model/juno_sw_comparision")
end
```

![](../figures/model/juno_sw_comparision.png)