### 2025-05-03 notebooks/21_model.qmd
refactor: better plot

# MHD Model and Juno solar wind parameters comparison

@wilsonSolarWindProperties2018

```{julia}
using Discontinuity
using Discontinuity: compute_params!, Alfven_current, inertial_length, gradient_current, uless, plasma_beta
using Juno: produce_or_load_juno
using Juno.Data: read_sw_params_csv
```

```{julia}
using DataFrames, DataFramesMeta
using Unitful

jno_df = produce_or_load_juno()
jno_sw_params = read_sw_params_csv(; add_unit=true)
```

```{julia}
using FlexiJoins
using Dates
using Unitful: μ0

jno_df_with_params = innerjoin(
    (jno_df, jno_sw_params),
    by_pred(x -> x.time .. (x.time + Minute(1)), ∋, :time);
)

@chain jno_df_with_params begin
    compute_params!
    @transform! begin
        :V_n_cross_sw = @. abs(first(:n_cross) * :V_KMPS)
        :V_n_mva_sw = @. abs(first(:n_mva) * :V_KMPS)
        :d_i_sw = inertial_length.(:N_PROTONS_CC, Unitful.q, Unitful.mp)
        :J_A_sw = Alfven_current.(:B_mag, :N_PROTONS_CC)
        :β_sw = plasma_beta.(:T_PROTONS_EV, :N_PROTONS_CC, :B_mag)
    end
    @transform! @astable begin
        :L_n_cross_sw = :V_n_cross_sw .* :duration
        :L_n_cross_sw_norm = @. :L_n_cross_sw / :d_i_sw |> NoUnits
        :J_m_max_cross_sw = @. abs(gradient_current(:grad, :V_n_cross_sw))
        :J_m_max_cross_sw_norm = @. :J_m_max_cross_sw / :J_A_sw |> NoUnits

        :L_n_mva_sw = :V_n_mva_sw .* :duration
        :L_n_mva_sw_norm = @. :L_n_mva_sw / :d_i_sw |> NoUnits
        :J_m_max_mva_sw = @. abs(gradient_current(:grad, :V_n_mva_sw))
        :J_m_max_mva_sw_norm = @. :J_m_max_mva_sw / :J_A_sw |> NoUnits
    end
end
```

```{julia}
using CairoMakie
import CairoMakie.Makie.SpecApi as S
using UnitfulLatexify: latexroundunitlabel as lru

uless_plot!(ax, x, y; kw...) = scatter!(ax, uless.(x), uless.(y); kw...)

let df = jno_df_with_params, xlabel = "Model", ylabel = "Juno", axis = (; xscale=log10, yscale=log10), plot = (; alpha=0.3)
    f = Figure()
    ax1 = Axis(f[1, 1]; title=lru("V", u"km/s"), ylabel)
    ax2 = Axis(f[1, 2]; title=lru("n", u"cm^-3"), axis...)

    ax_T = Axis(f[2, 1]; title=lru("T", u"eV"), axis...)
    ax_β = Axis(f[2, 2]; title=lru("β", u""), axis...)

    ax3 = Axis(f[4, 1]; title=lru("L", u"km"), ylabel, axis...)
    ax4 = Axis(f[4, 2]; title=lru("J", u"nA/m^2"), axis...)
    ax5 = Axis(f[5, 1]; title=L"L/d_i", xlabel, ylabel, axis...)
    ax6 = Axis(f[5, 2]; title=L"J/J_{Alfven}", xlabel, axis...)

    @with df begin
        uless_plot!(ax1, :V_mag, :V_KMPS)
        uless_plot!(ax2, :n, :N_PROTONS_CC)

        p1 = uless_plot!(ax3, :L_n_cross, :L_n_cross_sw, label="Cross"; plot...)
        p2 = uless_plot!(ax3, :L_n_mva, :L_n_mva_sw, label="MVA"; plot...)
        Legend(f[3, :], [p1, p2], ["Cross", "MVA"])

        uless_plot!(ax_T, :T, :T_PROTONS_EV)
        uless_plot!(ax_β, :β, :β)

        uless_plot!(ax4, :J_m_max_cross, :J_m_max_cross_sw, label="Cross"; plot...)
        uless_plot!(ax4, :J_m_max_mva, :J_m_max_mva_sw, label="MVA"; plot...)
        uless_plot!(ax5, :L_n_cross_norm, :L_n_cross_sw_norm; plot...)
        uless_plot!(ax5, :L_n_mva_norm, :L_n_mva_sw_norm; plot...)
        uless_plot!(ax6, :J_m_max_cross_norm, :J_m_max_cross_sw_norm; plot...)
        uless_plot!(ax6, :J_m_max_mva_norm, :J_m_max_mva_sw_norm; plot...)
    end

    color = :red
    linestyle = :dash

    for ax in (ax1,)
        ablines!(ax, 0, 1; color, linestyle)
    end

    lines!(ax2, exp10.(-1:0), exp10.(-1:0); color, linestyle)
    lines!(ax3, exp10.([2, 4.5]), exp10.([2, 4.5]); color, linestyle)
    lines!(ax4, exp10.(-2:1), exp10.(-2:1); color, linestyle)
    lines!(ax5, exp10.(-1:2), exp10.(-1:2); color, linestyle)
    lines!(ax6, exp10.(-2:1), exp10.(-2:1); color, linestyle)

    limits!(ax1, 360, 550, 360, 550)

    easy_save("model/juno_sw_comparision")
end
```

![](../figures/model/juno_sw_comparision.png)